---
title: "Lab Assignment 1"
author: "Zheshuo and Geming"
date: "2025-10-05"
output: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
---

# ASSIGNMENT 1: GEORGIA HOUSE PRICE

## 0. Before Starting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Clear the plots
rm(list=ls()) # Clear the environment
set.seed(123) # Set a seed to reproduce outcomes
```

We start by loading essential libraries and importing the data.

```{r}
# Installing dependencies
#install.packages("lubridate")
```

```{r include = FALSE}
# Loading libraries
library(moments)
library(car)
library(chemometrics)
library(mice)
library(missMDA)
library(FactoMineR)
library(corrplot)
library(VIM)
library(dplyr)
library(ggplot2)
library(AER)
library(MASS)
library(dplyr)
library(lubridate)
library(stringr)
library(readr)
library(tidyr)
library(tidyselect)
library(effects)
```

```{r}
# Importing the main data of the assignment
georgia <- read.csv("RealEstate_Georgia.csv")
```

The dataset contains 6168 housing ads from different cities in different counties of the Georgia state, there are posts from 2020 to 2021 but majority of data are for 2021. For a fair analysis, the houses that are posted in 2021 are taken into account.

```{r}
# Get the dimension of the dataset
dim(georgia) # 6168 obs + 39 variables
```

```{r}
# Show the name of columns
names(georgia)
```

```{r}
# Check the structure of the dataset
str(georgia)
```

```{r}
# Use summary function to get a fast understanding of the dataset
summary(georgia)
```

```{r}
# Extract years from the column datePostedString
georgia$datePostedString <- as.Date(georgia$datePostedString, format = "%Y-%m-%d")
georgia$datePostedString <- as.numeric(format(georgia$datePostedString, "%Y"))
```

## 1. Explanatory Data Analysis

In this section, we perform a crucial step in any data science project, especially when preparing data for modeling. A comprehensive descriptive analysis is carried out for each variable, emphasizing their most relevant characteristics. We proceed with a column-by-column examination of the dataset, followed by the generation of a data quality report. Subsequently, we prepare the data for the modeling phase by handling missing values through imputation techniques, removing non-informative variables, and creating new derived features where appropriate.

```{r}
# The function we need for EDA processes
# Use given Q1 & Q3 to compute indexes of outliers
calculate_outliers <- function(x) {
  # Check for regular outliers
  col_name <- deparse(substitute(x))
  Boxplot(x, main = "Boxplot", ylab = col_name)
  outliers <- boxplot.stats(x)$out

  # Compute quartiles and IQR
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1

  # Define thresholds
  lower_mild <- Q1 - 1.5 * IQR
  upper_mild <- Q3 + 1.5 * IQR
  lower_extreme <- Q1 - 3 * IQR
  upper_extreme <- Q3 + 3 * IQR

  # Identify mild and extreme outliers
  mild_idx <- which((x < lower_mild | x > upper_mild))
  extreme_idx <- which((x < lower_extreme | x > upper_extreme))
  
  # Count
  num_mild <- length(mild_idx)
  num_extreme <- length(extreme_idx)

  abline(h = c(lower_mild, upper_mild), col = "blue", lwd = 2)
  abline(h = c(lower_extreme, upper_extreme), col = "red", lwd = 2)

  # Return the results as a list
  return(list(
    num_mild = num_mild,
    num_extreme = num_extreme,
    mild_indexes = mild_idx,
    extreme_indexes = extreme_idx
  ))
}
```

**Variable 1: X.1**

Variable X.1 is a numerical variable with no missing values.
A quick inspection reveals that it serves as a unique identifier for each observation, since the number of unique values equals the total number of rows.
Therefore, this variable does not contain meaningful numerical information for statistical analysis. Normality and outlier tests were not performed, as the variable is an ID rather than a continuous numeric feature. It contains no missing values thus imputation is not needed. This variable does not contain any outliers. Interestingly, the maximum value (6468) in the data does not match the number of columns (6168), maybe these ids belong to the observations of the original dataset where rental houses from other states are also stored.

```{r}
# Summary of the column
summary(georgia$X.1)
# Any missing values
sum(is.na(georgia$X.1))
# First five elements
georgia$X.1[1:5]
# Number of unique values
unique_counts<-length(unique(georgia$X.1))
# Verify if all values are unique
unique_counts == nrow(georgia)
```

**Variable 2: X**

X is a numerical variable without missing values. It is very similar to the previous variable in all aspects. Although the range of its values is twice of the previous one ([1,6428] vs [25,13803]).

```{r}
# Summary of the column
summary(georgia$X)
# Any missing values
sum(is.na(georgia$X))
# First five elements
georgia$X[1:5]
# Number of unique values
unique_counts<-length(unique(georgia$X))
# Verify if all values are unique
unique_counts == nrow(georgia)

```

**Variable 3: id**

This categorical variable is suppose to contain the unique identifier of each row.
This variable does not have missing values. However, the number of unique values (5515 levels) do not match with the number of rows (6168 rows), so we can't really use it as an identifier in that regard. Nonetheless, there is another possibility that the dataset contains duplicated rows. And it happens that our idea has hit the bull's eye, it is true that the rows with the same ids are duplicated, most of their variables are the same with exception of the columns X.1, X and countyId. However, we also find out that not all rows with the same id are equal in terms of other variables, there are some which are likely to have errors in variables like garageSpaces, hasGarage and so on. But only 25 ids out of 653 duplicated ids show that pattern. Moreover, in the examples we present below, all sensible observations seem to be the last one, so in the later data cleaning section, we will keep only those observations.

```{r}
# Summary of the column
summary(georgia$id)
# Any missing values
sum(is.na(georgia$id))
# First five elements
georgia$id[1:5]
# Number of unique values
length(unique(georgia$id))

# Get duplicated ids
non_unique_ids <- georgia$id[duplicated(georgia$id)]
length(non_unique_ids)

# Observations with the same id are equal in most of the variables
which(georgia$id == non_unique_ids[1])

# obs 19 vs obs 5718
georgia[c(2975, 4502),]

# Let's try another pair
which(georgia$id == non_unique_ids[2])

# obs 19 vs obs 5718
georgia[c(1212, 4505),]

# Ignore the columsn X.1, X and countyId
ignore_cols <- c("X.1", "X", "countyId")
compare_cols <- setdiff(names(georgia), c(ignore_cols))

# For each duplicated id, check if all its rows are identical
same_rows <- sapply(non_unique_ids, function(idv) {
  sub <- georgia[georgia$id == idv, compare_cols, drop = FALSE]
  # All rows identical if only 1 unique row remains after removing duplicates
  nrow(unique(sub)) == 1
})

# Are all duplicates identical?
all(same_rows) # No

# Show two duplicates with different values in those variables
singular_ids <- names(same_rows)[which(same_rows == FALSE)]
georgia[which(georgia$id == singular_ids[1]),]
georgia[which(georgia$id == singular_ids[2]),]

# Number of singular ids
length(singular_ids)
```

**Variable 4: stateId**

This numerical variable contains the id of the state of the rental house in USA. However, since we only have observations from Georgia state, all these ids are the same (Georgia state's id is 16 for all rows). And it has no missing values.

```{r}
# Summary of the column
summary(georgia$stateId)
# Any missing values
sum(is.na(georgia$stateId))
# First five elements
georgia$stateId[1:5]
# Number of unique values
length(unique(georgia$stateId))

# Check frequency
table(georgia$stateId)

```

**Variable 5: countyId**

This numerical variable contains the id of counties in Georgia state. It has no missing values, and the number of unique values is slightly lower than the number of rows, which suggest some of the counties are used more than once. The data is not normally distributed and there is no presence of outliers.

```{r}
# Summary of the column
summary(georgia$countyId)
# Any missing values
sum(is.na(georgia$countyId))
# First five elements
georgia$countyId[1:5]

# Number of unique values
length(unique(georgia$countyId))

hist(georgia$countyId, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$countyId), sd(georgia$countyId)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sci <- sample(georgia$countyId, 5000)
shapiro.test(sci)

# Boxplot to detect outliers
calculate_outliers(georgia$countyId)
```

**Variable 6: cityId**

This numerical variable represents the unique identifiers for cities in the state of Georgia. The cityId values range from 0 to 397,383, with 374 unique values. Notably, there are 340 occurrences of the value 0, which might indicate missing data.

```{r}
# Summary of the column
out_ci <- summary(georgia$cityId)
out_ci
# Any missing values
sum(is.na(georgia$cityId))
# First five elements
georgia$cityId[1:5]
# Number of unique values
length(unique(georgia$cityId))

# Check frequency
table(georgia$cityId)

```

**Variable 7: country**

This categorical variable contains the name of the country where the rental houses are found. Each house is geographically located in Georgia, a state in the USA, so its the same value for all the rows. This variable has no missing values.

```{r}
# Summary of the column
summary(georgia$country)
# Any missing values
sum(is.na(georgia$country))
# First five elements
georgia$country[1:5]
# Number of unique values
length(unique(georgia$country))
# Check frequency
table(georgia$country)

```

**Variable 8: datePostedString**

This variable was previously a categorical variable which contains the date when the rental house was published. We performed a conversion at the beginning to keep only the year. This variable has no missing values and most of the houses were posted on 2021, only a small number of them are posted 2020 (47).

```{r}
# Summary of the column
summary(georgia$datePostedString)
# Any missing values
sum(is.na(georgia$datePostedString))
# First five elements
georgia$datePostedString[1:5]
# Number of unique values
length(unique(georgia$datePostedString))

length(which(georgia$datePostedString != 2021))

plot(factor(georgia$datePostedString))
```

**Variable 9: is_bankOwned**

The variable is_bankOwned is a boolean variable that indicates whether the house is offered by a bank. This variable has no missing values and almost no house is owned by a bank, with exception of the observation 158.

```{r}
# Summary of the column
summary(georgia$is_bankOwned)
# Any missing values
sum(is.na(georgia$is_bankOwned))
# First five elements
georgia$is_bankOwned[1:5]
# Number of unique values
length(unique(georgia$is_bankOwned))

length(which(georgia$is_bankOwned != 0))
# Houses owned by a bank
which(georgia$is_bankOwned != 0)

plot(factor(georgia$is_bankOwned))
```

**Variable 10: is_forAuction**

The variable is_forAuction is almost identical to the previous column in terms of labels. It is also a boolean variable that indicates whether the house belongs to an auction. This variable has no missing values and no house is auctioned, with exception of the observation 3946.

```{r}
# Summary of the column
summary(georgia$is_forAuction)
# Any missing values
sum(is.na(georgia$is_forAuction))
# First five elements
georgia$is_forAuction[1:5]
# Number of unique values
length(unique(georgia$is_forAuction))

length(which(georgia$is_forAuction != 0))
# Houses belong to an auction
which(georgia$is_forAuction != 0)

plot(factor(georgia$is_forAuction))
```

**Variable 11: event**

The variable event is a categorical variable with 5 levels that indicates the sales type of each rental house. This variable has no missing values, around 5000 observations have "Listed for sale" and around 1000 have "Price change" in this variable. The rest of the labels have very low presence in the data.

```{r}

# Any missing values
sum(is.na(georgia$event))
# First five elements
georgia$event[1:5]
# Number of unique values
length(unique(georgia$event))

# Convert 'event' to factor for better handling as categorical variable
georgia$event <- factor(georgia$event)
# Summary of the column
summary(georgia$event)
plot(factor(georgia$event))
```

**Variable 12: time**

The time variable is a numerical variable representing the timestamp when each observation was collected. This variable has no missing values and contains 216 unique timestamps. The data was collected after 2021.

```{r}
# Summary of the column
out_ct <- summary(georgia$time)
out_ct
# Any missing values
sum(is.na(georgia$time))
# First five elements
georgia$time[1:5]
# Number of unique values
length(unique(georgia$time))

hist(georgia$time, breaks = 20, freq = F)
# Check for any duplicate timestamps (indicating possible data duplication)
sum(duplicated(georgia$time))  # Count how many duplicate timestamps exist

# convert time stamp to date
dates <- as.POSIXct(georgia$time / 1000, origin = "1970-01-01", tz = "UTC")
dates

```

**Variable 13: price**

The variable price is a numerical variable indicating the price of the rental houses. It ranges from 6,250 USD to 1,300,000 USD. This variable has no missing values and there are 1227 unique values. It's distribution resembles that of a normal distribution, but the Shapiro test says the otherwise. Moreover, there are 43 extreme outliers in the data.

```{r}
# Summary of the column
out_cp <- summary(georgia$price)
out_cp
# Any missing values
sum(is.na(georgia$price))
# First five elements
georgia$price[1:5]
# Number of unique values
length(unique(georgia$price))

hist(georgia$price, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$price), sd(georgia$price)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sp <- sample(georgia$price, 5000)
shapiro.test(sp)

# Detect outlier
out <- calculate_outliers(georgia$price)

# Print number of extreme outliers
out$num_extreme

```

**Variable 14: pricePerSquareFoot**

The variable pricePerSquareFoot is a numerical variable that, as its name indicates, it stores the price per square foot of the rental houses. It ranges from 4.0 sqft to 205.00 sqft. This variable has no missing values and there are 440 unique values. It's not normally distributed and there are some considerably high values that might be errors while collecting the data. Those values are most likely extreme outliers in the data, which are 56.

```{r}
# Summary of the column
out_cppsf <- summary(georgia$pricePerSquareFoot)
out_cppsf
# Any missing values
sum(is.na(georgia$pricePerSquareFoot))
# First five elements
georgia$pricePerSquareFoot[1:5]
# Number of unique values
length(unique(georgia$pricePerSquareFoot))

hist(georgia$pricePerSquareFoot, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$pricePerSquareFoot), sd(georgia$pricePerSquareFoot)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sppsf <- sample(georgia$pricePerSquareFoot, 5000)
shapiro.test(sppsf)

# Detect outlier
out <- calculate_outliers(georgia$pricePerSquareFoot)

# Print number of extreme outliers
out$num_extreme
```

**Variable 15: city**

The variable city is a categorical variable indicating the name of the city where the houses are located in. There are no missing data in the data and it has 383 unique values in this column. The most frequent city that appears in this data is Atlanta, followed by Marietta, Augusta, Columbus and Macon.

```{r}
# Summary of the column
summary(georgia$city)
# Any missing values
sum(is.na(georgia$city))
# First five elements
georgia$city[1:5]
# Number of unique values
length(unique(georgia$city))

plot(factor(georgia$city))
sort(table(georgia$city), decreasing = TRUE)[1:5]
```

**Variable 16: state**

This categorical variable contains the name of the state where the rental houses are found. Each house is geographically located in Georgia, so in this field we only have one value, which is Georgia, represented by GA. This variable has no missing values.

```{r}
# Summary of the column
summary(georgia$state)
# Any missing values
sum(is.na(georgia$state))
# First five elements
georgia$state[1:5]
# Number of unique values
length(unique(georgia$state))
```

**Variable 17: yearBuilt**

The variable yearBuilt is a numerical variable indicating the year when the house was built. There are 159 unique values. It's proved by the Shapiro test that it is not normally distributed and there are 16 extreme outliers. Those outliers are irreal, the maximum value of the data is 9999 which might indicate the presence of missing data in the dataset.

```{r}
# Summary of the column
out_cyb <- summary(georgia$yearBuilt)
out_cyb
# Any missing values
sum(is.na(georgia$yearBuilt))
# First five elements
georgia$yearBuilt[1:5]
# Number of unique values
length(unique(georgia$yearBuilt))

hist(georgia$yearBuilt, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$yearBuilt), sd(georgia$yearBuilt)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
syb <- sample(georgia$yearBuilt, 5000)
shapiro.test(syb)

# Detect outlier
out <- calculate_outliers(georgia$yearBuilt)

# Print number of extreme outliers
out$num_extreme
```

**Variable 18: streetAddress**

This categorical variable contains the street address of the rental houses. There are no missing values and there are 5514 levels, meaning some houses are geographically located in the same street.

```{r}
# Summary of the column
summary(georgia$streetAddress)
# Any missing values
sum(is.na(georgia$streetAddress))
# First five elements
georgia$streetAddress[1:5]
# Number of unique values
length(unique(georgia$streetAddress))
```

**Variable 19: zipcode**

Zipcode is a numerical variable containing the zipcode of each rental house. There are no missing values and there are 446 unique numerical values. The data is not normally distributed (confirmed by the Shapiro test). In this case we do not have any extreme outliers, but we do have 331 mild outliers.

```{r}
# Summary of the column
out_czc <- summary(georgia$zipcode)
out_czc
# Any missing values
sum(is.na(georgia$zipcode))
# First five elements
georgia$zipcode[1:5]
# Number of unique values
length(unique(georgia$zipcode))

hist(georgia$zipcode, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$zipcode), sd(georgia$zipcode)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
szc <- sample(georgia$zipcode, 5000)
shapiro.test(szc)

# Detect outlier
out <- calculate_outliers(georgia$zipcode)

# Print number of extreme outliers
out$num_mild
```

**Variable 20: longitude**

The numerical variable longitude contains the longitude of each of the rental houses. There are no missing values and there are 5437 unique numerical values. The data is not normally distributed (confirmed by the Shapiro test) and ranges from -85.57 to -80.97. In addition, there are 97 extreme outliers.

```{r}
# Summary of the column
out_clg <- summary(georgia$longitude)
out_clg
# Any missing values
sum(is.na(georgia$longitude))
# First five elements
georgia$longitude[1:5]
# Number of unique values
length(unique(georgia$longitude))

hist(georgia$longitude, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$longitude), sd(georgia$longitude)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
slg <- sample(georgia$longitude, 5000)
shapiro.test(slg)

# Detect outlier
out <- calculate_outliers(georgia$longitude)

# Print number of extreme outliers
out$num_extreme
```

**Variable 21: latitude**

The numerical variable latitude contains the latitude of each of the rental houses. There are no missing values and there are 5455 unique numerical values. The data is not normally distributed (confirmed by the Shapiro test) and ranges from 30.73 to 34.99. In addition, there are 82 extreme outliers.

```{r}
# Summary of the column
out_clt <- summary(georgia$latitude)
out_clt
# Any missing values
sum(is.na(georgia$latitude))
# First five elements
georgia$latitude[1:5]
# Number of unique values
length(unique(georgia$latitude))

hist(georgia$latitude, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$latitude), sd(georgia$latitude)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
slt <- sample(georgia$latitude, 5000)
shapiro.test(slt)

# Detect outlier
out <- calculate_outliers(georgia$latitude)

# Print number of extreme outliers
out$num_extreme
```

**Variable 22: hasBadGeocode**

This boolean variable indicates whether a row has bad geocode. Geocode allows we to take each observation's information and create a map of their locations. There are no missing values in the data, but there are only 0s thoughout the column, in other words, there are no observation with bad geocode, since we can determine the location of each house using the longitude and latitude columns in the dataset.

```{r}
# Summary of the column
summary(georgia$hasBadGeocode)
# Any missing values
sum(is.na(georgia$hasBadGeocode))
# First five elements
georgia$hasBadGeocode[1:5]
# Number of unique values
length(unique(georgia$hasBadGeocode))

hist(georgia$hasBadGeocode, breaks = 20, freq = F)
```

**Variable 23: description**

This categorical variable contains a textual description provided by the sellers of the rental houses. There are no missing values in the data, but not all descriptions are unique (unique values = 5429), there are some rows that share the same descriptions. This reminds us of the analysis we did with the variable 3: id. We found out that there are some observations with the same observations are equal in most of the variables, but there are some that do not follow that pattern. Maybe the sellers were so lazy that they copied and pasted one of the description they had on their hand on similar rental houses' posts.

```{r}
# Summary of the column
summary(georgia$description)
# Any missing values
sum(is.na(georgia$description))
# First five elements
georgia$description[1]
# Number of unique values
length(unique(georgia$description))

# Get duplicated descriptions
duplicated_descriptions <- georgia[duplicated(georgia$description) | duplicated(georgia$description, fromLast = TRUE), ]

# Some observations with the same description are equal in most of the variables
which(georgia$description == duplicated_descriptions$description[1])

# obs 19 vs obs 5718
georgia[c(19, 5718),]

# Not all observations with the same description are equal
which(georgia$description == duplicated_descriptions$description[6])

# obs 60 vs obs 174
georgia[c(60, 174),]
```

**Variable 24: currency**

This categorical variable contains the currency used in each of the observations. However, since we only have observations from Georgia state, the currency in each row is USA. And it has no missing values.

```{r}
# Summary of the column
summary(georgia$currency)
# Any missing values
sum(is.na(georgia$currency))
# First five elements
georgia$currency[1:5]
# Number of unique values
length(unique(georgia$currency))
```

**Variable 25: livingArea**

This numerical variable contains the living are of the rental houses in sqft. There are no missing values and it's not normally distributed. Although we said that it has no missing values, there is an observation where the livingArea is 1 (minimum value), this could be an error or missing value. There are in total 2601 unique values in the data. In addition, 9 extreme outliers are found, each of them are above 7000 sqft.

```{r}
# Summary of the column
out_cla <- summary(georgia$livingArea)
out_cla
# Observation with livingArea = 1
which(georgia$livingArea < 100) # 1282
# Any missing values
sum(is.na(georgia$livingArea))
# First five elements
georgia$livingArea[1:5]
# Number of unique values
length(unique(georgia$livingArea))

hist(georgia$livingArea, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$livingArea), sd(georgia$livingArea)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sla <- sample(georgia$livingArea, 5000)
shapiro.test(sla)

# Detect outlier
out <- calculate_outliers(georgia$livingArea)

# Print number of extreme outliers
out$num_extreme
```

**Variable 26: livingAreaValue**

This numerical variable contains exactly the same values as the previous one, they are actually the same columns with a different column name.

```{r}
# Whether these two columns are equal
all(georgia$livingArea == georgia$livingAreaValue)
```
**Variable 27: bathrooms**

This numerical variable indicates the number of bathrooms each one of the rental houses has. There are no missing values (NAs) and it's not normally distributed. Thre are 11 unique values and 35 extreme outliers (> 6 bathrooms). Some of the observations have 0 bathrooms, these must be an error, so we will deal with them later with imputation.

```{r}
# Summary of the column
out_cb <- summary(georgia$bathrooms)
out_cb
# Any missing values
sum(is.na(georgia$bathrooms))
# First five elements
georgia$bathrooms[1:5]
# Number of unique values
length(unique(georgia$bathrooms))

hist(georgia$bathrooms, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$bathrooms), sd(georgia$bathrooms)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sb <- sample(georgia$bathrooms, 5000)
shapiro.test(sb)

# Detect outlier
out <- calculate_outliers(georgia$bathrooms)

# Print number of extreme outliers
out$num_extreme
```

**Variable 28: bedrooms**

This numerical variable indicates the number of bedrooms each one of the rental houses has. There are no missing values (NAs) and it's not normally distributed. There are 12 unique values and 18 extreme outliers (> 7 bedrooms). Similar to the previous variable, some of the observations have 0 bedrooms, so we will also deal with them later with imputation.

```{r}
# Summary of the column
out_cbd <- summary(georgia$bedrooms)
out_cbd
# Any missing values
sum(is.na(georgia$bedrooms))
# First five elements
georgia$bedrooms[1:5]
# Number of unique values
length(unique(georgia$bedrooms))

hist(georgia$bedrooms, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$bedrooms), sd(georgia$bedrooms)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sbd <- sample(georgia$bedrooms, 5000)
shapiro.test(sbd)

# Detect outlier
out <- calculate_outliers(georgia$bedrooms)

# Print number of extreme outliers
out$num_extreme
```

**Variable 29: buildingArea**

This numerical variable contains exactly the same values as the variables 24 and 25 (livingArea and livingAreaValue), they are actually the same columns with a different column name.

```{r}
# Whether these two columns are equal
all(georgia$livingAreaValue == georgia$buildingArea)
```

**Variable 30: parking**

The variable parking is a boolean variable that indicates whether the house offers parking. This variable has no missing values. The number of houses with parking (4711) exceeds the number of those which have not (1457).

```{r}
# Summary of the column
summary(georgia$parking)
# Any missing values
sum(is.na(georgia$parking))
# First five elements
georgia$parking[1:5]
# Number of unique values
length(unique(georgia$parking))

plot(factor(georgia$parking))
```

**Variable 31: garageSpaces**

The variable garageSpaces is a numerical variable that indicates the number of garage spaces in the house. This variable has no missing values and there are 8 unique values from 0 to 7. It's not normally distributed and there are no extreme outliers, but it has 5 mild outliers (> 5 garage spaces).

```{r}
# Summary of the column
out_cgs <- summary(georgia$garageSpaces)
out_cgs
# Any missing values
sum(is.na(georgia$garageSpaces))
# First five elements
georgia$garageSpaces[1:5]
# Number of unique values
length(unique(georgia$garageSpaces))

hist(georgia$garageSpaces, breaks = 20, freq = F)
curve(dnorm(x, mean(georgia$garageSpaces), sd(georgia$garageSpaces)), add = T)

# Shapiro test only accepts sample sizes between 3 and 5000
sgs <- sample(georgia$garageSpaces, 5000)
shapiro.test(sgs)

# Detect outlier
out <- calculate_outliers(georgia$garageSpaces)

# Print number of extreme outliers
out$num_mild
```

**Variable 32: hasGarage**

The variable hasGarage is a boolean variable that indicates whether the house has a garage. This variable has no missing values. The number of houses with a garage (3821) exceeds the number of those which have not (2347).

```{r}
# Summary of the column
summary(georgia$hasGarage)
# Any missing values
sum(is.na(georgia$hasGarage))
# First five elements
georgia$hasGarage[1:5]
# Number of unique values
length(unique(georgia$hasGarage))

plot(factor(georgia$hasGarage))

# Is there a house where the variable hasGarage is 0 and there are more than 0 garageSpaces?
which(georgia$garageSpaces > 0 & georgia$hasGarage == 0)
```

**Variable 33: levels**

The variable levels is a categorical variable that indicates the levels of the buildings. This variable has no missing values and it has 35 unique values, although the level "0" might indicate missing values in this variable. However, this variable is a bit messy, there are a lot of repeated labels, they may be written differently, but in fact, they have the same meaning. For instance, the label "3 Story" matches with "Tri-level".

```{r}
# Summary of the column
summary(georgia$levels)
# Any missing values
sum(is.na(georgia$levels))
# First five elements
georgia$levels[1:5]
# Number of unique values
length(unique(georgia$levels))

# Show all levels
unique(georgia$levels)
```

**Variable 34: pool**

The variable pool is a boolean variable that indicates whether the house has a pool This variable has no missing values. Only 438 observations out of 6168 have a pool.

```{r}
# Summary of the column
summary(georgia$pool)
# Any missing values
sum(is.na(georgia$pool))
# First five elements
georgia$pool[1:5]
# Number of unique values
length(unique(georgia$pool))

plot(factor(georgia$pool))

# Number of houses with pool
length(which(georgia$pool == 1))
```

**Variable 35: spa**

The variable spa is a boolean variable that indicates whether the house has spa This variable has no missing values. Only 644 observations out of 6168 have spa.

```{r}
# Summary of the column
summary(georgia$spa)
# Any missing values
sum(is.na(georgia$spa))
# First five elements
georgia$spa[1:5]
# Number of unique values
length(unique(georgia$spa))

plot(factor(georgia$spa))

# Number of houses with spa
length(which(georgia$spa == 1))
```

**Variable 36: isNewConstruction**

The variable isNewConstruction is a boolean variable that indicates whether a rental house is a new building. There are no missing values and only 388 observations out of 6168 are new.

```{r}
# Summary of the column
summary(georgia$isNewConstruction)
# Any missing values
sum(is.na(georgia$isNewConstruction))
# First five elements
georgia$isNewConstruction[1:5]
# Number of unique values
length(unique(georgia$isNewConstruction))

plot(factor(georgia$isNewConstruction))

# Number of new constructed houses
length(which(georgia$isNewConstruction == 1))
```

**Variable 37: hasPetsAllowed**

The variable hasPetsAllowed is a boolean variable that indicates whether it is allowed to have pets in the rental houses. There are no missing values and only 51 houses out of 6168 allow pets.

```{r}
# Summary of the column
summary(georgia$hasPetsAllowed)
# Any missing values
sum(is.na(georgia$hasPetsAllowed))
# First five elements
georgia$hasPetsAllowed[1:5]
# Number of unique values
length(unique(georgia$hasPetsAllowed))

plot(factor(georgia$hasPetsAllowed), )

# Number of pets allowing houses
length(which(georgia$hasPetsAllowed == 1))
```

**Variable 38: homeType**

The variable homeType is a categorical variable that indicates the type of the rental houses. This variable has 5 unique values and there are no missing values. The type "SINGLE_FAMILY" is predominant in the data, the rest of the labels have low presence throughout the data.

```{r}
# Summary of the column
summary(georgia$homeType)
# Any missing values
sum(is.na(georgia$homeType))
# First five elements
georgia$homeType[1:5]
# Number of unique values
length(unique(georgia$homeType))

plot(factor(georgia$homeType))
```

**Variable 39: county**

The variable county is a categorical variable that indicates the county in Georgia state where the building is located. This variable has 129 unique values and there are no missing values.

```{r}
# Summary of the column
summary(georgia$county)
# Any missing values
sum(is.na(georgia$county))
# First five elements
georgia$county[1:5]
# Number of unique values
length(unique(georgia$county))
```
## 2. Data preparation

In this section we will prepare the data in order to facilitate the later analysis. In the previous section we found different problems of each variable, so here we are going to either apply different cleaning functions to the columns or just remove them.

```{r}
# There are duplicated rows/obs in the dataset, we can use the column id to remove them
duplicated_rows <- duplicated(georgia$id, fromLast = TRUE) # Find duplicated rows from the bottom
georgia_clean <- georgia[!duplicated_rows, ] # Remove duplicated observations
```

```{r}
# We only want to keep observations from 2021
georgia_clean <- georgia_clean[georgia_clean$datePostedString == 2021,]
```

```{r}
# The column levels is messy and not consistent, since it has both numeric and categorical labels

# We are going to perform an explicit mapping for the 35 observed raw values into 14 categories + NA, which will be imputed after analyzing the data quality
mapping <- c(
  "One and One Half"                          = "1.5",
  "One"                                       = "1",
  "Two"                                       = "2",
  "Three Or More"                             = "3+",
  "Multi/Split"                               = "multi-split",
  "Multi/Split-Split Level"                   = "multi-split",
  "0"                                         = NA, # This value will be imputed
  "One-Two Story Foyer"                       = "multi-foyer",
  "Tri Level"                                 = "tri-level",
  "2.5 Story"                                 = "2.5",
  "Manufactured Home 1 Story"                 = "manufactured",
  "Two-Split Foyer"                           = "split-foyer",
  "One and One Half-Split Level"              = "multi-split",
  "Split Foyer"                               = "split-foyer",
  "One and One Half-Two"                      = "multi",
  "Two-Two Story Foyer"                       = "multi",
  "One and One Half-Multi/Split"              = "multi-split",
  "Three Or More-Two Story Foyer"             = "3+",
  "One-Two"                                   = "multi",
  "One-One and One Half"                      = "multi",
  "One-Manufactured Home 1 Story"             = "manufactured",
  "Multi/Split-Two"                           = "multi-split",
  "Two-Multi/Split"                           = "multi-split",
  "One-Other"                                 = "other",
  "Two-Three Or More"                         = "multi",
  "One-One and One Half-Two"                  = "multi",
  "Three Or More-Split Level-Tri-Level"       = "3+",
  "Two-Foyer - 2 Story"                       = "2",
  "Other-See Remarks"                         = "other-remarks",
  "Split Level"                               = "split",
  "3 Story"                                   = "3+",
  "Two-Split Level"                           = "split",
  "One-Mobile Home 1 Story"                   = "manufactured",
  "Tri Level-Split Level"                     = "tri-level",
  "Tri-Level"                                 = "tri-level"
)

# Apply mapping
georgia_clean$levels_mapped <- mapping[georgia_clean$levels]

```

```{r}
# Regarding the columns city/cityId and county/countyId, we noticed that they do not have the same number of unique variables. By the way, the column cityId has some missing values where the cityId equals 0.
length(unique(georgia$cityId)) == length(unique(georgia$city))

# It's because of that, in the following cell we will remove the columns cityId and countyId and retain city and county

```

```{r}
# The following columns just don't provide any useful information

# Columns to remove (we chose livingArea instead of pricePerSquareFoot)
cols_to_remove <- c(
  "X.1", "X", "id", "stateId", "country", "time", "cityId", "datePostedString",
  "state", "streetAddress", "hasBadGeocode", "description", 
  "currency", "livingAreaValue", "buildingArea", "hasPetsAllowed", 
  "is_bankOwned", "is_forAuction", "countyId", "pricePerSquareFoot", "levels"
)

# Remove them from the dataset
georgia_clean <- georgia_clean[, setdiff(names(georgia_clean), cols_to_remove)]

# Check remaining columns
names(georgia_clean)
```

```{r}
# Rental houses built in the future are not possible
georgia_clean[georgia_clean$yearBuilt > 2025,"yearBuilt"] <- NA
# Rental houses with 0 bedrooms or bathrooms are not acceptable
georgia_clean[georgia_clean$bedrooms == 0,"bedrooms"] <- NA
georgia_clean[georgia_clean$bathrooms == 0,"bathrooms"] <- NA
# In case hasGarage == 0 and garageSpaces > 0
georgia_clean[georgia_clean$hasGarage == 0 & georgia_clean$garageSpaces > 0,"hasGarage"] <- NA
georgia_clean[georgia_clean$hasGarage == 0 & georgia_clean$garageSpaces > 0,"garageSpaces"] <- NA
# In livingArea there is an observation with livingArea = 1
georgia_clean[georgia_clean$livingArea < 100,"livingArea"] <- NA

# These are preliminary extreme outliers we found, later we will assign NAs to each of the outliers found in these numerical variables using Boxplot

# As discussed in the exploratory data analysis, these missing values are likely the result of later updates to the rental houses. This could explain the presence of duplicated rows, or they might simply come from errors introduced during data collection or upload.
```

```{r}
# Removing trailing blanks and converting every category to lower case
georgia_clean <- georgia_clean %>%
  mutate(across(where(~ is.character(.) | is.factor(.)),
                ~ tolower(trimws(as.character(.)))))
# Doing this can reduce the number of unique values in categorical variables. For example, in the variable city we had originally 383 factors, but in the filtered dataset we have now 371 factors.
length(unique(georgia$city)) # Original dataset
length(unique(georgia_clean$city)) # Filtered dataset

```


```{r}
summary(georgia_clean)
```

As we can see from the summary statistics, the data generally makes sense and follows the expected rules for the housing field. Prices range from 6,250 USD to 1.3 million USD, which is plausible given the variability of the Georgia housing market. Structural variables like livingArea, bedrooms, and bathrooms mostly fall within reasonable ranges, though the maximums (10 bathrooms, 14 bedrooms) suggest a few outliers. yearBuilt spans 1800 to 2022, reflecting a mix of older and newer houses. Geographic variables (latitude, longitude, zipcode) lie within Georgia’s bounds. Logical patterns are generally consistent, for example, homes marked with hasGarage = 1 usually have a positive number of garageSpaces, and only a small fraction of properties are labeled as new constructions. Overall, the data supports the working theory that property characteristics such as size and age influence price, while occasional anomalies are likely due to minor data entry or collection errors rather than systemic issues in the dataset.

## 3. Data quality reports

**Variables**

Up to this points, all variables in the dataset have been explored. In terms of missingness, some of the variables have exaggerated values or error values, those could be considered as missing values. If we rank the numeric variables in terms of number of outliers, the following list is obtained, starting with the most outliers: longitude (97 extreme outliers), latitude (82 extreme outliers), pricePerSquareFoot (56 extreme outliers), price (43 extreme outliers), bathrooms (35 extreme outliers), bedrooms (18 extreme outliers), yearBuilt (16 extreme outliers), livingArea (9 extreme outliers, we are ignoring livingAreaValue and buildingArea, since they have the same values), zipcode (331 mild outliers) and garageSpaces (5 mild outliers).
 
**Univariate Analisis**

Now we will study the univariate outliers of the aforementioned variables. Firstly, the number of univariate outliers per individual are counted and added in a new variable called ‘univ_outl_count’. Looking at the individuals with the most univariate outliers (3) it can be concluded that they are all costly and have unusually high number of bathrooms and bedrooms.

The correlation matrix shows a positive correlation between price, bathrooms, bedrooms, livingArea and garageSpaces. This makes sense, as larger houses typically command higher prices due to the increased living space and associated amenities. Interestingly, the number of outliers appears to be negatively correlated with latitude, suggesting that properties located at lower latitudes tend to have more outliers.

```{r}
# longitude
georgia_clean$univ_outl_count <- 0

# Calculate outliers
out <- calculate_outliers(georgia_clean$longitude)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "longitude"] <- NA
```

```{r}
# latitude

# Calculate outliers
out <- calculate_outliers(georgia_clean$latitude)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "latitude"] <- NA
```

```{r}
# price

# Calculate outliers
out <- calculate_outliers(georgia_clean$price)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
# see what happen in extreme, because this is target value, we cannot do imputation.
extreme_rows <- georgia_clean[out$extreme_indexes, , drop = FALSE]
extreme_rows

# we can see extreme outliers are luxury house, we can remove them, now we add na to label them.
georgia_clean[out$extreme_indexes, "price"] <- NA


```

```{r}
# bathrooms

# Calculate outliers
out <- calculate_outliers(georgia_clean$bathrooms)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "bathrooms"] <- NA
```

```{r}
# bedrooms

# Calculate outliers
out <- calculate_outliers(georgia_clean$bedrooms)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "bedrooms"] <- NA
```

```{r}
# yearBuilt

# Calculate outliers
out <- calculate_outliers(georgia_clean$yearBuilt)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "yearBuilt"] <- NA
```

```{r}
# livingArea

# Calculate outliers
out <- calculate_outliers(georgia_clean$livingArea)
# Update the column univ_outl_count
georgia_clean[out$extreme_indexes, "univ_outl_count"] = georgia_clean[out$extreme_indexes, "univ_outl_count"] + 1
# Add NAs
georgia_clean[out$extreme_indexes, "livingArea"] <- NA
```

```{r}
# zipcode

# Calculate outliers
out <- calculate_outliers(georgia_clean$zipcode)
# Update the column univ_outl_count
georgia_clean[out$mild_indexes, "univ_outl_count"] = georgia_clean[out$mild_indexes, "univ_outl_count"] + 1
# Add NAs (here we use mild outliers, since we don't have extreme outliers)
georgia_clean[out$mild_indexes, "zipcode"] <- NA
```

```{r}
# garageSpaces

# Calculate outliers
out <- calculate_outliers(georgia_clean$garageSpaces)
# Update the column univ_outl_count
georgia_clean[out$mild_indexes, "univ_outl_count"] = georgia_clean[out$mild_indexes, "univ_outl_count"] + 1
# Add NAs (here we use mild outliers, since we don't have extreme outliers)
georgia_clean[out$mild_indexes, "garageSpaces"] <- NA
```

```{r}
# There are observations with 3 outliers in those columns
max(georgia_clean$univ_outl_count)
# Print those individuals
georgia_clean[which(georgia_clean$univ_outl_count == 3),]
```

```{r}
# We are going to a create a correlation matrix of those columns
georgia_interest = georgia_clean[,c("longitude", "latitude", "price", "bathrooms", "bedrooms", "yearBuilt", "livingArea", "zipcode", "garageSpaces", "univ_outl_count")]
# Use pairwise.complete.obs to ignore NAs in each pair of variables
cor_outl = cor(georgia_interest, use = "pairwise.complete.obs")
par(mfrow=c(1,1))
corrplot(cor_outl, method = 'number', number.cex = 0.5)
```

**Data Imputation**

Here we will use imputePCA function to impute numerical data and imputeMCA for categorical variables.

```{r}
# first we remove all na's in target price to avoid imputation to target value.
georgia_clean <- drop_na(georgia_clean,"price")
# We will use imputePCA to impute numerical variables
res.misMDA <- imputePCA(georgia_clean[, c(2,4:16)]) # Choosing numerical variables
# Get the completed dataset
georgia_clean[, c(2,4:16)] <- res.misMDA$completeObs
```

```{r}
# We will use imputeMCA to impute categorical variables
res.misMDA <- imputeMCA(georgia_clean[,c(1,3,17:19)]) # Choosing categorical variables
# Get the completed dataset
georgia_clean[, c(1,3,17:19)] <- res.misMDA$completeObs
```

```{r}
summary(georgia_clean) # The distribution of the imputed dataset seems to align the previous one
```

**Multivariate Analisis**

Moutlier is applied on the numerical variables to find multivariate outliers. With the garageSpaces variable included, however, the method returns a singular matrix. Therefore this variable is excluded from the calculation. A very mild threshold of 0.025 % is chosen as signficance level because is already returns a significant amount of outliers. This makes up around 7% of the total amount of instances. It is chosen to delete these outliers from the data set for the rest of the project. 

```{r}
# Use Moutlier that applies Mahalanobis distances to detect multivariate outliers
res.out = Moutlier(georgia_clean[,c("longitude", "latitude", "price", "bathrooms", "bedrooms", "yearBuilt", "livingArea", "zipcode")], quantile = 0.975, col="green") # Condidence level of 97.5%
```

```{r}
length(which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff))
 )/nrow(georgia_clean)
```

```{r}
# Plot all observations
par(mfrow=c(1,1))
plot(res.out$md, res.out$rd, xlab = "Mahalanobis Distance", ylab = "Robust Distance") 
abline(h=res.out$cutoff, col="red") 
abline(v=res.out$cutoff, col="red") 

```

```{r}
# Summary of only multivariate outliers
summary(georgia_clean[which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff)),])
```

```{r}
# Summary of the original dataset
summary(georgia_clean)
```

```{r}
# Remove those observations from the dataset
georgia_clean = georgia_clean[-which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff)),] 
```

After imputation and elimination of outliers, we can start doing different analysis of the data and creating our own linear models. However, some variables such as bathrooms, bedrooms, parking, garageSpaces, hasGarage, pool, spa and isNewConstruction are either binary or categorized numerical variables. Therefore, in the next cell we are going to convert them into factors.

```{r}
# From binary variables to factors
georgia_clean$parking <- factor(georgia_clean$parking, labels = c("No-Parking", "Yes-Parking"))
georgia_clean$hasGarage <- factor(georgia_clean$hasGarage, labels = c("No-Garage", "Yes-Garage"))
georgia_clean$pool <- factor(georgia_clean$pool, labels = c("No-Pool", "Yes-Pool"))
georgia_clean$spa <- factor(georgia_clean$spa, labels = c("No-Spa", "Yes-Spa"))
georgia_clean$isNewConstruction <- factor(georgia_clean$isNewConstruction, labels = c("Old", "New"))
```

After inspecting the variables bathrooms, bedrooms and garageSpaces, we decided to create three factors for all these three variables.

```{r}
# Inspecting
quantile(georgia_clean$bathrooms, probs=seq(0,1,0.1))
quantile(georgia_clean$bedrooms, probs=seq(0,1,0.1))
quantile(georgia_clean$garageSpaces, probs=seq(0,1,0.1))

# From categorized numeric variables to factors
georgia_clean$bathrooms[which(georgia_clean$bathrooms <= 2)] <- 0
georgia_clean$bathrooms[which(georgia_clean$bathrooms > 2 & georgia_clean$bathrooms <= 4)] <- 1
georgia_clean$bathrooms[which(georgia_clean$bathrooms > 4)] <- 2
georgia_clean$bathrooms <- factor(georgia_clean$bathrooms, labels = c("Little-Bathrooms", "Many-Bathrooms", "Excessive-Bathrooms"))

georgia_clean$bedrooms[which(georgia_clean$bedrooms <= 2)] <- 0
georgia_clean$bedrooms[which(georgia_clean$bedrooms > 2 & georgia_clean$bedrooms <= 4)] <- 1
georgia_clean$bedrooms[which(georgia_clean$bedrooms > 4)] <- 2
georgia_clean$bedrooms <- factor(georgia_clean$bedrooms, labels = c("Little-Bedrooms", "Many-Bedrooms", "Excessive-Bedrooms"))

georgia_clean$garageSpaces[which(georgia_clean$garageSpaces == 0)] <- 0
georgia_clean$garageSpaces[which(georgia_clean$garageSpaces > 0 & georgia_clean$garageSpaces <= 2)] <- 1
georgia_clean$garageSpaces[which(georgia_clean$garageSpaces > 2)] <- 2
georgia_clean$garageSpaces <- factor(georgia_clean$garageSpaces, labels = c("No-Spaces", "Has-Spaces", "Many-Spaces"))
```

```{r}
summary(georgia_clean)
```


**Profiling**

### Determine if the target variable (price) has an acceptably normal distribution. Address test to discard serial correlation. 

Now, we'd like to determine if the response variable (price) has an acceptably normal distribution. In addition, we used the acf() function to test for autocorrelation. The results indicate that the variable is not normally distributed and exhibits significant autocorrelation. The first few lags show high and positive autocorrelation, then it decreases gradually, indicating that values of price are correlated with their past values (the data are not independent). Therefore, the dataset is randomized to remove serial correlation before further analysis.

```{r}
# Plots the autocorrelation function to detect serial correlation in the data.
acf(georgia_clean$price)
# Serial correlation means that observations in our data are related across time, rather than being completely random or independent.
```

```{r}
# We can choose at most 5000 samples for shapiro test
ll <- sample(1:nrow(georgia_clean),nrow(georgia_clean), size = 5000) # Shuffling indexes
georgia_clean <- georgia_clean[ll,] # Shuffling data
acf(georgia_clean$price) 
```

```{r}
# We can choose at most 5000 samples for shapiro test
shapiro.test(georgia_clean$price[1:5000]) # Use the first 5000 obs

# H0: The price variable is normally distributes
# H1: The price variable is NOT normally distributes

# Since the p-value < 0.05, we reject H0.

# The variable georgia$price does not follow a normal distribution, it significantly deviates from normality.
```

```{r}
# This checks whether a sample comes from a specific theoretical distribution, in this case, a normal distribution with the same mean and standard deviation as our data.
ks.test(georgia_clean$price, 'pnorm', mean(georgia_clean$price), sd(georgia_clean$price))

# H0: georgia$price follows a normal distribution with the given mean and standard deviation
# H1: georgia$price does not follow that normal distribution.

# Because the p-value < 0.05, we reject H0.

# The K-S test confirms that the distribution of georgia$price significantly deviates from normality, meaning the data is not normally distributed.
```

```{r}
# Plot a histogram of price with a density curve
ggplot(data = georgia_clean, aes(x = price, y = after_stat(density))) + 
  geom_histogram(breaks = seq(0, max(georgia_clean$price), by = 10000),
                 col = "lightblue",
                 fill = "steelblue") +
  geom_density(lwd = 1, col = "red") +
  labs(title = "Histogram of Price with Density Curve",
       x = "Price",
       y = "Density")
```

### Indicate by exploration of the data which are apparently the variables most associated with the target variable Price (use only the indicated variables). 

To do this, the condes function of the FactoMiner package is used, which for the 
numeric target variable 'price' calculates the correlation of each of the quantitative variables and the coefficient of determination (R^2) for the qualitative variables, together with a p-value for significance.

For the quantitative variables, it is clear that livingArea, yearBuilt and latitude are positively correlated with price. This is logical, as larger and more recently built properties typically command higher prices, and latitude may capture differences in neighborhoods or proximity to desirable areas. Longitude and zipcode are negatively correlated with price (r < -0.10, p ~ 0), which is also expected, reflecting that certain geographic areas, often located further from city centers or in less affluent regions, tend to have lower property values.

For the qualitative variables, it is clear that city and bathrooms explain the most variance in the price variable (R^2 = 0.39, p ~ 0), followed by county (R^2 = 0.24), bedrooms (R^2 = 0.18), levels_mapped (R^2 = 0.18), garageSpaces (R^2 = 0.16) and hasGarage (R^2 = 0.15), all highly significant. This is expected, as geographic location and property layout strongly influence housing prices. Although, theoretically, bedrooms should be expected to have a stronger correlation with price, in this case, bathrooms appear more decisive, possibly because properties in Georgia do not always need many bedrooms. Other categorical variables, such as spa, parking, pool and homeType, have a much smaller effect (R^2 ~ 0.01), indicating only a minor association with price.

For the different categories in the categorical variables, houses with excessive bathrooms and bedrooms have a mean price much higher than the grand mean price. Some city categories, like milton, bishop, white plains, marble hill, and brooks, also show mean prices well above the grand mean, meaning properties in those areas tend to be more expensive. On the other hand, houses with fewer bathrooms or bedrooms have mean prices lower than the grand mean, along with other categories like no-Garage, no-Spa, or 1 level. Overall, this makes sense, since bigger houses with more rooms usually cost more, while simpler homes or those with fewer features tend to be cheaper.

```{r}
# Displays the summary of variables most related to the 2th column (price), showing how each numeric variable contributes or associates with it.
res.con = condes(georgia_clean,2) # The column corresponding to price

# Quantitative variables
res.con$quanti
# The p.value here assesses statistical significance of the correlation between each variable and price

# Qualitative variables
res.con$quali
# The coefficient of determination measures the proportion of variance in price explained by each categorical variable

# Focusing on categories
res.con$category
```

### Define a polytomic factor f.yearBuilt for the building year of a house according to its quartiles and argue if the average price depends on the age of the building.

```{r}
# Create the polytomic factor by yearBuilt quartiles
yrs <- georgia_clean$yearBuilt
qs <- quantile(yrs, probs = c(0, 0.25, 0.5, 0.75, 1))
georgia_clean <- georgia_clean %>%
  mutate(f.yearBuilt = cut(yearBuilt,
                           breaks = qs,
                           include.lowest = TRUE,
                           labels = c("Old","MidOld","New","Newest")))
```

Given the evidence below, we argue that the price is dependent on the level of yearBuilt. Firstly, in the boxplots it can be observed that the older the building (Q1) the lower the median price of the establishment This is also seen clearly in the distribution plot below. Secondly, the Wilcoxon test shows that the medians across the yearBuilt levels are not equal, indicating a statistically significant relationship between building age and price.

```{r}
# Price distribution differs across building-age groups
ggplot(georgia_clean, aes(x=price, fill=f.yearBuilt)) + geom_density(alpha=.5)

# Boxplots
ggplot(georgia_clean, aes(y=price, fill=f.yearBuilt)) + geom_boxplot(alpha=0.5)
```

We perform pairwise wilcoxon test to check for similar medians. Looking at the boxplot and distribution plot we can already expect that these are not going to be similar. The result of the wilcoxon test indicates our hypothesis was right that there is a clear difference between the medians of the different quartiles.

```{r}
# The following non-parametric test checks whether the median prices between each pair of building-age groups differ significantly.
pairwise.wilcox.test(georgia_clean$price, georgia_clean$f.yearBuilt)

# H0: Median(price_1) = Median(price_2)
# H1: Median(price_1) != Median(price_2)

# Since all p-values are extremely small (p < 0.05), we reject H0 for all pairs, meaning the median prices differ significantly between at least some of the f.yearBuilt categories.
```

### Interpretation of rental house price according to the building year factor and the home type

First, box plots of the home type and building year factor are presented to show that there is variability in the means of both factors, meaning that they both could have significant influence on the rental house price. Next to the boxplots, wilcoxon and fligner tests are performed to assess if the medians and means of the categories in the factors are statistically different and if the variances between the categories in the factors are statistically different.

```{r}
# Summarizes the frequency of each homeType category
summary(georgia_clean$homeType)

# Creates boxplots of price for each homeType
boxplot(price ~ homeType, data = georgia_clean)

# Displays boxplots of price grouped by building-age quartiles
boxplot(price ~ f.yearBuilt, data = georgia_clean)

# Tests if median prices differ significantly between building-age groups
pairwise.wilcox.test(georgia_clean$price, georgia_clean$f.yearBuilt)
# There is a significant difference in mean between each of the groups

# Tests if median prices differ significantly between home types
pairwise.wilcox.test(georgia_clean$price, georgia_clean$homeType, conf.level=0.99)
# The H0 of mean homogeneity is rejected for the groups townhouse and single_family
# Therefore, we can say that townhouse and single family are homogeneous in terms
# of group mean

# Tests if variance in Price variable differ significantly between building-age groups
fligner.test(georgia_clean$price, georgia_clean$f.yearBuilt)
# There is a significant difference in variance between each of the groups

# Tests if variance in Price variable differ significantly between home types
fligner.test(georgia_clean$price, georgia_clean$homeType)
# There is a significant difference in variance between each of the groups
```

From the interaction plots, it can be observed that the building year factor has a large influence on the price, and that although smaller, the home types also influence the price. Moreover, it’s important to note that for the home type multi_family, there are no observations for the MidOld and Newest building year categories, which may limit the interpretation of interaction effects. It seems that newer buildings consistently yield higher prices across most home types, while older properties tend to cluster around lower price range.

```{r}
# Note that some combinations do not exist, so they won't appear in the plot
table(georgia_clean$homeType, georgia_clean$f.yearBuilt)
# For home types MidOld, New (only 1) and Newest we do not have observations
 
par(mfrow=c(1,1)) 
# Create an interaction plot of price by homeType and f.yearBuilt
interaction.plot(georgia_clean$homeType, georgia_clean$f.yearBuilt, georgia_clean$price)
# Create an interaction plot of price by f.yearBuilt and homeType
par(mfrow=c(1,1)) 
interaction.plot(georgia_clean$f.yearBuilt, georgia_clean$homeType, georgia_clean$price)
```

## 4. Customizing linear models to predict the Price variable

We start our first model using all the available numerical variables.

From the *Residuals vs Fitted plot*, we observe that most residuals are centered around 0, which is good. However, there is a noticeable pattern: the residuals appear more dispersed for lower fitted values and become tighter as fitted values increase. This suggests that cheaper houses show more variability than more expensive ones, indicating possible heteroscedasticity.

The *Normal Q-Q plot* shows that the standardized residuals follow the theoretical normal line quite well in the middle range but deviate in both tails, more strongly in the upper tail. This indicates that the residuals are not perfectly normally distributed, with potential outliers or skewness affecting the fit.

The *Scale-Location plot* confirms the observation of increasing spread of residuals, as the red trend line shows a slight upward slope, reinforcing the presence of heteroscedasticity.

The *Residuals vs Leverage plot* shows that most observations have low leverage, and no points appear to have both high leverage and large standardized residuals, suggesting no influential outliers with undue impact on the model.

Finally, since the Variance Inflation Factor (VIF) values are around 1, multicollinearity is not a concern, and all predictors can be retained.

To better satisfy the linear regression assumptions, particularly homoscedasticity and normality, we will apply appropriate transformations to the data (such as log or BoxCox transformations).

```{r}
# Define the baseline model based on relevant numrical variables
m1_0 <- lm(price ~ livingArea + yearBuilt + latitude, data = georgia_clean)
# Summary of the model
summary(m1_0) # Multiple R-squared: 0.5729; All coefficients are significant
plot(m1_0)
vif(m1_0)
```

Using the Box-Cox transformation, it can be observed that a transformation is appropriate. Since the estimated λ (lambda) is close to 0, (between 0 and 1) a logarithmic transformation of the response variable seems suitable.

```{r}
# Do I need to apply some transformation to my output?
boxcox(price ~ livingArea + yearBuilt + latitude, data = georgia_clean)
```

In the following cell we assess whether we need to apply some transformations to the explanatory variables. The variable livingArea shows a MLE of lambda of 0.07663 (close to zero) that suggests a log transform is appropriate. The variable yearBuilt's MLE of lambda is far from 1, but since it is statistically insignificant, no transformation is needed. The variable latitude's MLE of lambda is also far from 1 and highly significant (p < 2e-16), but extreme lambda suggests instability, so we are going to standardize this variable first. After applying a minmax scale to the latitude, we get a MLE of lambda of -2.6976 and this suggests that we should apply a quadratic transformation to it.

```{r}
# Do I need to apply some transformations to my explanatory variables?
boxTidwell(log(price) ~ livingArea + yearBuilt + latitude, data = georgia_clean)

# Standardization needed for the variable latitude (min-max scale)
georgia_clean$latitude.scaled <- ( georgia_clean$latitude -
                                   min(georgia_clean$latitude)
                                 ) / 
                                 ( max(georgia_clean$latitude) -
                                   min(georgia_clean$latitude)
                                 ) + 1 # values between 1 and 2

# Check the boxTidwell function again
boxTidwell(log(price) ~ log(livingArea) + yearBuilt + latitude.scaled, data = georgia_clean)

# Check the boxTidwell function after applying the quadratic transformation
m1_1 <- lm(log(price) ~ log(livingArea) + yearBuilt + poly(latitude.scaled,2), data = georgia_clean)
# Summary of the model
summary(m1_1) # Multiple R-squared: 0.5835; All coefficients are significant
plot(m1_1)
vif(m1_1)

residualPlots(m1_1)
# The line is roughly horizontal and near 0
# And the spread of residuals does not change much across x-values (homoscedasticity)
avPlots(m1_1, id=list(method=list(cooks.distance(m1_1),"x"), n = 5))
marginalModelPlots(m1_1)
# The last two plots show slightly inclined lines around 0 and that means that
# the model’s residuals still have a small amount of patterns unexplained
# by our model. In other words, the relationship between the model and the
# target variable price may be not fully linear, even after our current
# transformations.

```

```{r}
# Here we can see the log effect of livingArea and the quadratic effect for latitude
plot(allEffects(m1_1, transformation=list(link=log, inverse=exp)))
```

```{r}
# In this cell we investigate if it is really worthy to apply a quadratic
# transformation to the latitude variable
# In order to use anove on two nested models, we apply log transformations to
# the variables price and livingArea to the model m1_0
m1_0t <- lm(log(price) ~ log(livingArea) + yearBuilt + latitude, data = georgia_clean) # linear effect of latitude
m1_1 <- lm(log(price) ~ log(livingArea) + yearBuilt + poly(latitude.scaled,2), data = georgia_clean) # quadratic effect of latitude

anova(m1_1, m1_0t) # The model m1_1t contains m1_0t, these two are nested
# H0: The shorter model is the best
# Since the p-value is near 0, we reject the null hypothesis thus we opt for
# the complex model

# We can also check the AIC values of each model
AIC(m1_1) # 4904.767 -> AIC value is lower, we opt for it
AIC(m1_0t) # 5160.026

```

```{r}
# We have seen 4 methods for detecting infueltial observations:

# Identify observations with large studentized residuals (> 3), 
# which may indicate outliers in terms of the response variable.
llres <- which(abs(rstudent(m1_1)) > 3); length(llres)
# Identify observations with high leverage values, those that have 
# an unusually strong influence on the fitted values. 
# The threshold 3*p/n (where p = number of coefficients) is a common rule of thumb.
llhat <- which(hatvalues(m1_1) > 3 * length(coef(m1_1)) / nrow(georgia_clean)); length(llhat)
# Plot Cook’s distance to visually identify influential points. 
# The Boxplot function highlights observations with unusually high influence.
llcoo <- Boxplot(cooks.distance(m1_1), id = list(n=10, labels = seq(nrow(georgia_clean)))); length(llcoo)
# Identify observations with Cook’s distance greater than 4/(n - p), 
# which is another rule of thumb to flag influential cases 
# that strongly affect the regression coefficients.
llCH <- which(cooks.distance(m1_1) > 4 / (nrow(georgia_clean) - length(coef(m1_1)))); length(llCH)
# Combine all influential observation indices (row positions)
influential_ids <- unique(c(llres, llhat, llcoo, llCH))

# But we opted for finding influential observations using the ids returned by incluencePlot
# function, since it retuns only the most extreme ones, instead of finding them in a 
# systematic way.

# Check influential observations and remove them from the dataset
influential <- influencePlot(m1_1, pch=19,id=list(method="noteworthy",n=3))
df_used <- model.frame(m1_1)   
idx  <- as.integer(match(rownames(influential), rownames(df_used)))
# Check if all visually flagged influential points are also in the combined list
all(idx %in% influential_ids)
# Remove them from the dataset
georgia_clean <- georgia_clean[-idx,]
```

```{r}
# Let's create another model based on the cleaned dataset
m1_2 <- lm(log(price) ~ log(livingArea) + yearBuilt + poly(latitude.scaled,2), data = georgia_clean)
```

Previously we created a discretized variable from the quartiles of the numeric variable yearBuilt and now we want to check whether it is better to add this variable as a numeric or discretized variable.

```{r}
# Check proportions
table(georgia_clean$f.yearBuilt)

# Do note that when we add a categorical variable as an explanatory variable to a linear model, the number of coefficients we obtain associated to this variable is the number of categories - 1. 
# The category that is not on the estimation is the baseline category (comes from baseline reparametrization).
# The coefficient associated to this category is assumed to be equal to (alpha_1=0)

# Build a similar model to m1_2 using discretized varaible f.yearBuilt
m1_3 <- lm(log(price) ~ log(livingArea) + f.yearBuilt + poly(latitude.scaled,2), data = georgia_clean)
summary(m1_3)
# Estimate: effects over the baseline category (the baseline here is 'Old')
# If we go from MidOld to Newest, the values will increase in 0.13452 (0.16112 - 0.02660)

# We need to compare between two models:
AIC(m1_2, m1_3) # As the Akaike value is lower in m1_2, we will use the numerical one
# They are not nested, they have the same number of explanatory variables,
# but they do not have the same number of categories
```

Next, we will add main effects of factors. We will only include those relevant factors found using previously in the profiling section through the condes() function from the FactoMineR library.

```{r}
# Add relevant categorized numeric variables
m2_0 <- lm(log(price) ~ log(livingArea) + 
                        yearBuilt + 
                        poly(latitude.scaled,2) +
                        bathrooms +
                        city +
                        county +
                        bedrooms + 
                        levels_mapped + 
                        garageSpaces + 
                        hasGarage +
                        spa +
                        parking + 
                        pool + 
                        homeType, data = georgia_clean)
# We only print the Multiple R-squared value,
# since city and county have too many categories
summary(m2_0)$r.squared
plot(m2_0)
#step(m2_0)
```
 
Adding categorical variables improved performance, but the models became more prone to overfitting. Since the city and county fields have high cardinality (many levels), and stepwise selection returned several NA coefficients, this might also be caused by factors with 0 occurrences in the dataset. Therefore, we decided to reduce their cardinality by keeping only those cities and counties with occurrences above the median.

```{r}
# Convert city and county factors to characters to remove empty levels
georgia_clean$city <- as.character(georgia_clean$city)
georgia_clean$county <- as.character(georgia_clean$county)

# Keep cities with occurrences above the median
tab_city <- table(georgia_clean$city)
keep_cities <- names(tab_city[tab_city > median(tab_city)]) # median = 7

# Keep counties with occurrences above the median
tab_county <- table(georgia_clean$county)
keep_counties <- names(tab_county[tab_county > median(tab_county)]) # meadian = 17

# Set cities and counties with ocurrences below the median as Other
georgia_clean$city_grp <- ifelse(
  georgia_clean$city %in% keep_cities,
  georgia_clean$city, "Other"
)
georgia_clean$county_grp <- ifelse(
  georgia_clean$county %in% keep_counties,
  georgia_clean$county,
  "Other"
)

# Store the modified city and county variables as factors
georgia_clean$city_grp <- factor(georgia_clean$city_grp,
                                   levels = c(keep_cities, "Other"))
georgia_clean$county_grp <- factor(georgia_clean$county_grp,
                                   levels = c(keep_counties, "Other"))

# Change back the previously altered city and county variables to factors
georgia_clean$city <- factor(georgia_clean$city)
georgia_clean$county <- factor(georgia_clean$county)

```

```{r}
# Check the same model again using the filtered city and county variables
m2_1 <- lm(log(price) ~ log(livingArea) + 
                        yearBuilt + 
                        poly(latitude.scaled,2) +
                        bathrooms +
                        city_grp +
                        county_grp +
                        bedrooms + 
                        levels_mapped + 
                        garageSpaces + 
                        hasGarage +
                        spa +
                        parking + 
                        pool + 
                        homeType, data = georgia_clean)

# Print the Multiple R Squared value of the model
summary(m2_1)$r.squared
```

After reducing the cardinality of those factors and filtering the categories in the city and county variables, we will analyze the interactions between different numerical and categorical variables. We can start by performing an interaction analysis between the numerical variable log(livingArea) and the categorical variable bathrooms.

```{r}
# Interaction: log(livingArea):bathrooms
m3_0 <- lm(log(price) ~ log(livingArea) *
                        bathrooms +
                        yearBuilt + 
                        poly(latitude.scaled,2) +
                        city_grp +
                        county_grp +
                        bedrooms + 
                        levels_mapped + 
                        garageSpaces + 
                        hasGarage +
                        spa +
                        parking + 
                        pool + 
                        homeType, data = georgia_clean)

summary(m3_0) # Having Little bathrooms, the values increase in 0.797 units

# Does the effect (by default 0.797) of log(livingArea) depend on the number of bathrooms?.
# Does the slope change if we have Little, Mant or Excessive bathrooms?

# We use log(livingArea) * bathrooms
# We have two extra categories:
# If we have (baseline) log(livingArea):bathroomsLittle-Bathrooms -> log(livingArea): 0.797
# If we have (new) log(livingArea):bathroomsMany-Bathrooms -> log(livingArea): 0.796 + 0.704 = 1.5
# If we have (new) log(livingArea):bathroomsExcessive-Bathrooms -> log(livingArea): 0.796 + 2.88 = 3.676

# Plotting the effects of the interaction
plot(Effect(c("bathrooms", "livingArea"), m3_0,
            transformation=list(link=log, inverse=exp)),
     axes=list(y=list(lab="price")))

# Testing the model with the previous model
anova(m2_1, m3_0) # low p-value suggests that the model m3_0 is better

# Which variable are relevant?
Anova(m3_0) # Anova checks the significance of all categorical variables
# It seems that all variables we have so far are significant
#step(m3_0) # step function also suggests that the current setup is suitable, we do
# not need to remove variables
AIC(m3_0) # 2279.166

```

Now, we add all interactions to the model and then use step-wise selection to obtain the best specification.

```{r}
# Adding all interactions
m3_1 <- lm(log(price) ~ log(livingArea) * (
                        bathrooms +
                        yearBuilt + 
                        poly(latitude.scaled,2) +
                        city_grp +
                        county_grp +
                        bedrooms + 
                        levels_mapped + 
                        garageSpaces + 
                        hasGarage +
                        spa +
                        parking + 
                        pool + 
                        homeType), data = georgia_clean)

# Summary
summary(m3_1)$r.squared
# Step-wise selection of interations
#step(m3_1)

# Let's also check if there is multicollinearity presence in the model
vif(m3_1, type = "predictor") # compute VIF for each original predictor
```

Looking at the output of the step() function, we see that the interactions log(livingArea):city_grp and log(livingArea):garageSpaces are the only redundancies, so we will update our model by excluding these two interactions from m3_1. From the table generated by vif(), we observe many -Inf or NaN values in the GVIF, indicating perfect collinearity caused by interactions or factors with many levels. This collinearity is highly likely due to the aforementioned interactions.

```{r}
# Excluding those interactions from the previous one
# After running again the step() function it suggests us to remove log(livingArea):parking
m4_0 <- lm(log(price) ~ log(livingArea) * (
                        bathrooms +
                        yearBuilt + 
                        poly(latitude.scaled,2) +
                        county_grp +
                        bedrooms + 
                        levels_mapped + 
                        hasGarage +
                        spa +
                        pool + 
                        homeType) +
                        city_grp +
                        garageSpaces +
                        parking, data = georgia_clean)

# Summary
summary(m4_0)$r.squared
# Check again if there are still unnecessary interactions in the model
#step(m4_0)
# Now none of the remaining interactions needs to be removed

# Let's also check is there is multicollinearity presence in the  best model
vif(m4_0, type = "predictor")
```

We can see that predictors with GVIF^(1/(2*Df)) values around 1 are acceptable, but those with much higher values indicate severe multicollinearity, meaning the variables are highly correlated and make the model’s interpretation more difficult.

We eliminated the perfect collinearity associated with latitude.scaled by removing its squared term and excluding it from interactions with log(livingArea). Likewise, we mitigated the severe multicollinearity involving levels_mapped by not including it in interactions with log(livingArea).

```{r}
# Handling severe multicollinearity caused by latitude.scaled and levels_mapped
m4_1 <- lm(log(price) ~ log(livingArea) * (
                        bathrooms +
                        yearBuilt +
                        county_grp +
                        bedrooms + 
                        hasGarage +
                        spa +
                        pool + 
                        homeType) +
                        city_grp +
                        garageSpaces +
                        parking +
                        levels_mapped +
                        latitude.scaled, data = georgia_clean)
# Summary
summary(m4_1)$r.squared
# Check again if there are still unnecessary interactions in the model
#step(m4_1)
# Let's also check is there is multicollinearity presence in the  best model
vif(m4_1, type = "predictor")

# After stepwise selection we find 2 innecesary interaction (with bedrooms and bathrooms)
m4_2= lm(log(price) ~ log(livingArea) *(   yearBuilt + county_grp + 
      hasGarage + spa + pool + homeType)+ bedrooms + bathrooms + city_grp + 
    garageSpaces + parking + levels_mapped + latitude.scaled , data = georgia_clean)
anova(m4_2, m4_1) 
vif(m4_2, type = "predictor")
summary(m4_2)$r.squared
```

Before validating the model, we need to check the influential values again to ensure that any remaining influence on the model is minimized. The influence plot below shows most cases with low leverage. Then, we remove the flagged cases and refit the model on the cleaned sample.
Comparing the Q–Q plots before and after removing these observations, there is a modest improvement in distributional fit.
```{r}
# Checking influential values using influencePlot
influential <- influencePlot(m4_2,pch=19,id=list(method="noteworthy",n=3))

rownames(influential)
df_used <- model.frame(m4_2)   
idx  <- as.integer(match(rownames(influential), rownames(df_used)))
df_used[idx,]
df_influential<-georgia_clean[idx,]
georgia_clean <- georgia_clean[-idx,]

m4_3= lm(log(price) ~ log(livingArea) *(   yearBuilt + county_grp + 
      hasGarage + spa + pool + homeType)+ bedrooms + bathrooms + city_grp + 
    garageSpaces + parking + levels_mapped + latitude.scaled , data = georgia_clean)
vif(m4_3, type = "predictor")
summary(m4_3)$r.squared
influencePlot(m4_3,pch=19,id=list(method="noteworthy",n=3))
plot(m4_2, which = 2)
plot(m4_3, which = 2)
# After removing these influential values we can see that the upper bound is 0.0159
```



Although severe multicollinearity (GVIF^(1/(2*Df)) > 10) still exists in the model, we decided to retain it as our best model and proceed with the validation process.

Residuals are approximately centered at zero and roughly symmetric. The LOESS smooth shows a slight arch with a dip at higher fitted values, suggesting a mild remaining systematic pattern.

The Q–Q plot shows points align with the 45° line in the center but diverge at both tails, indicating roughly symmetric, heavy-tailed residuals. Normality isn’t strictly satisfied, but the deviations are acceptable.
Compared with the initial specification, tail departures are smaller—especially on the right—and the central alignment is tighter, indicating a modest improvement in distributional fit.

The scale–location plot is nearly flat with a slight mid-range hump and no clear funnel, suggesting at most mild heteroskedasticity.

The residuals–leverage plot shows most observations with low leverage  and a small tail, without large residuals. No points appear beyond the usual Cook’s distance contours, indicating no clearly influential observations.


Theoretically, we shoud not use the dataset used for training for testing. But since it is specified in the pdf that we do not need to split the dataset intk train and test sets, we did not do thta at the beginning of the script.

On the training set, observed vs. predicted log(price) aligns closely with the 45° line, indicating good calibration and fit, dispersion is modest and slightly larger in the mid-range, consistent with mild heteroskedasticity.

Marginal plots indicate largely linear effects for log(livingArea) and yearBuilt, with a clearer hump for latitude.scaled. Overall fit is adequate.

In the model, price increases with living area (elasticity ≈ 0.386) and build year (~+1.86% per year), shows large amenity effects (garage ~+83%, spa ~+184%, pool ~−75%), small penalties for “many/excessive” bedrooms with small premiums for more bathrooms, and notable geography effects (several counties negative, many cities positive)

```{r}
# The best model we obtained
best_model <- m4_3

# Validation of the best model
par(mfrow=c(2,2))
plot(best_model)
par(mfrow=c(1,1))

library(ggplot2)
pred <- fitted(best_model); obs <- model.response(model.frame(best_model))  
ggplot(data.frame(pred, obs), aes(pred, obs)) +
  geom_point(alpha=.3) + geom_abline(linetype=2) +
  labs(x="Predicted log(price)", y="Observed log(price)")

library(effects)
plot(Effect(c("livingArea", "yearBuilt"), best_model), multiline=TRUE, ci.style="bars")
plot(Effect(c("livingArea", "spa"), best_model), multiline=TRUE, ci.style="bars")
plot(Effect(c("livingArea", "pool"), best_model), multiline=TRUE, ci.style="bars")
plot(Effect(c("livingArea", "hasGarage"), best_model), multiline=TRUE, ci.style="bars")
plot(Effect(c("livingArea", "homeType"), best_model), multiline=TRUE, ci.style="bars")

marginalModelPlots(best_model, id=list(n=5)) # Whether the model fits the data
# The lines should be as close as possible
summary(m4_3)
```
After completing validation, we conducted a broad influential-observation screen (studentized residuals, leverage, and Cook’s distance), removed the union of flagged cases, and refit the model.

Out of 4,981 observations, we flagged over 437 as outliers/influential; sensitivity checks showed material impact on coefficients and diagnostics, so we removed the union of these cases and re-estimated on the cleaned sample as a robustness check.

Comparison of the two sets of diagnostic plots

residuals vs fitted: The LOESS line is flatter and the end curvature is weaker, indicating less systematic structure in the residuals.

Q–Q plot: Tail departures shrink, especially on the right and the center now tracks the 45° line more closely.

scale–location: The mid-range peak has dropped and the spread is more even, leaving at most mild heteroskedasticity.

residuals–leverage: High-leverage cases now have smaller residuals, and no clear Cook’s-distance outliers are evident.

We conclude that the identified influential observations materially affected the model, after removing them, diagnostic fit improved and the specification became more robust.
```{r}
# Identify observations with large studentized residuals (> 3), 
# which may indicate outliers in terms of the response variable.
llres <- which(abs(rstudent(m4_3)) > 3); llres

# Identify observations with high leverage values, those that have 
# an unusually strong influence on the fitted values. 
# The threshold 3*p/n (where p = number of coefficients) is a common rule of thumb.
llhat <- which(hatvalues(m4_3) > 3 * length(coef(m4_3)) / nrow(georgia_clean)); llhat

# Plot Cook’s distance to visually identify influential points. 
# The Boxplot function highlights observations with unusually high influence.
llcoo <- Boxplot(cooks.distance(m4_3), id = list(labels = row.names(georgia_clean)))

# Identify observations with Cook’s distance greater than 4/(n - p), 
# which is another rule of thumb to flag influential cases 
# that strongly affect the regression coefficients.
llCH <- which(cooks.distance(m4_3) > 4 / (nrow(georgia_clean) - length(coef(m4_2)))); llCH

# Combine all influential observation indices (row positions)
influential_ids <- unique(c(llres, llhat, llcoo, llCH));length(influential_ids)
# Remove them from the dataset
georgia_clean_noinf <- georgia_clean[-as.numeric(influential_ids), ]

# Refit the model on the cleaned data
m4_4 <- update(m4_3, data = georgia_clean_noinf)
par(mfrow=c(2,2))
plot(m4_3)
plot(m4_4)
```



