---
title: "Lab_Assignment_2"
author: "Zheshuo and Geming"
date: "2025-11-10"
output: pdf_document
---


# ASSIGNMENT 2: HR Analytics - Job Change of Data Scientists 

## 0. Before Starting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Clear the plots
rm(list=ls()) # Clear the environment
```

We start by loading essential libraries and importing the data.

```{r}
# Installing dependencies
#install.packages("lubridate")
```

```{r include = FALSE}
# Loading libraries
library(moments)
library(car)
library(chemometrics)
library(mice)
library(missMDA)
library(FactoMineR)
library(corrplot)
library(VIM)
library(dplyr)
library(ggplot2)
library(AER)
library(MASS)
library(dplyr)
library(lubridate)
library(stringr)
library(readr)
library(tidyr)
library(tidyselect)
library(effects)
```

```{r}
# Importing the main data of the assignment
aug <- read.csv("aug_train.csv")
```

The dataset contains 19158 observations about different employees. Each row has 14 attributes/variables containing information related to demographics, education and experience of the employee.

```{r}
# Show the name of columns
names(aug)
```

```{r}
# Check the structure of the dataset
str(aug)
```

```{r}
# Here we convert the categorical variables into factors
aug$city <- factor(aug$city)
aug$gender <- factor(aug$gender)
aug$relevent_experience <- factor(aug$relevent_experience)
aug$enrolled_university <- factor(aug$enrolled_university)
aug$education_level <- factor(aug$education_level)
aug$major_discipline <- factor(aug$major_discipline)
aug$experience <- factor(aug$experience)
aug$company_size <- factor(aug$company_size)
aug$company_type <- factor(aug$company_type)
aug$last_new_job <- factor(aug$last_new_job)
```


```{r}
# Use summary function to get a fast understanding of the dataset
summary(aug)
```

```{r}
# Print the first 10 observations in the data
head(aug, 10)
```


## 1. Data Preparation

### 1.1. Univariate Descriptive Analysis

First of all, we can create a discretization version of variable for the the numerical variables city_development_index, training_hours and the target binary variable target. This can be done using the following cell. Moreover, the dataset has no apparent missing values (NAs), but we cannot discard the posibility of having them (missing values) as outliers in the dataset.

```{r}
# Creating the polytomic factors by city_development_index and training_hours quantiles
get_polytomic_factors <- function(x, variable, labels){
  # Compute quartile cut points
  qs <- quantile(variable, probs = c(0, 0.25, 0.5, 0.75, 1))
  
  # Derive a name for the new column (e.g. "f.city" if variable was aug$city)
  var_name <- deparse(substitute(variable))
  colname <- paste0("f.", substring(var_name, 5))
  
  # Create the factor column
  x[[colname]] <- cut(variable,
                      breaks = qs,
                      include.lowest = TRUE,
                      labels = labels)
  return(x)
}

# Get discretized numerical variables
aug <- get_polytomic_factors(aug, aug$city_development_index, labels = c("Low",
                                                                  "Med-Low",
                                                                  "Med-High",
                                                                  "High"))
aug <- get_polytomic_factors(aug, aug$training_hours, labels = c("Q1", "Q2", "Q3", "Q4"))
aug$target <- factor(aug$target, labels = c("NO-Job-Change", "YES-Job-Change"))

```

In the next cell we implement a function that calculates the univariate outliers of a given numeric variable in the dataframe object.

```{r}
# The function we need for EDA processes
# Use given Q1 & Q3 to compute indexes of outliers
calculate_outliers <- function(x, plot_yes = TRUE) {
  # Check for regular outliers
  col_name <- deparse(substitute(x))
  outliers <- boxplot.stats(x)$out

  # Compute quartiles and IQR
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1

  # Define thresholds
  lower_mild <- Q1 - 1.5 * IQR
  upper_mild <- Q3 + 1.5 * IQR
  lower_extreme <- Q1 - 3 * IQR
  upper_extreme <- Q3 + 3 * IQR

  # Identify mild and extreme outliers
  mild_idx <- which((x < lower_mild | x > upper_mild))
  extreme_idx <- which((x < lower_extreme | x > upper_extreme))
  
  # Count
  num_mild <- length(mild_idx)
  num_extreme <- length(extreme_idx)

  if (plot_yes) {
    Boxplot(x, main = "Boxplot", ylab = col_name)
    abline(h = c(lower_mild, upper_mild), col = "blue", lwd = 2)
    abline(h = c(lower_extreme, upper_extreme), col = "red", lwd = 2)
  }

  # Return the results as a list
  return(list(
    num_mild = num_mild,
    num_extreme = num_extreme,
    mild_indexes = mild_idx,
    extreme_indexes = extreme_idx
  ))
}

```

**Variable 1: enrollee_id**

Variable enrollee_id is a numerical variable with no missing values thus imputation is not needed. A quick inspection reveals that it serves as a unique identifier for each observation, since the number of unique values equals the total number of rows. Therefore, this variable does not contain meaningful numerical information for statistical analysis.

Normality and outlier tests were not performed, as the variable is an ID rather than a continuous numeric feature. Interestingly, the maximum value (33380) in the data does not match the number of columns (19158), which might mean that the dataset we have here is a subset of the original one. The variable enrollee_id ranges from 1 to 33380.

```{r}
# Summary of the column
summary(aug$enrollee_id)
# Any missing values?
sum(is.na(aug$enrollee_id))
# Number of unique values
unique_counts <- length(unique(aug$enrollee_id))
# Verify if all values are unique
unique_counts == nrow(aug)

```

**Variable 2: city**

The variable city is a categorical variable indicating the id of the city where the employees are located in. There are no missing data in the data and it has 123 unique values in this column. The most frequent city is city_103 with 4355 appearances and the lowest are city_127, city_133, city_167 and city_59 with 10 appearances.

```{r}
# Number of levels in the variable
length(levels(aug$city))
# Summary of the column
sort(summary(aug$city))
# Are there missing values?
sum(is.na(aug$city)) # No missing values

```

**Variable 3: city_development_index**

The variable city_development_index is a numerical variable indicating the development index of city. It ranges from 0.4480 to 0.9490. This variable has no missing values and there are 93 unique values. Based on histogram and Shapiro test we can consider it isn't normally distributed. Moreover, this variable doesn't have any extreme outlier but it has 17 mild outliers. In addition, looking at the barplot created using the discretized variable, we can say that the first two groups have similar number of observations. While the third group has the largest number of observations, the fourth group has the lowest.

```{r}
# Summary of the column
summary((aug$city_development_index))
# Any missing values?
sum(is.na(aug$city_development_index))
# Number of unique values
length(unique(aug$city_development_index))

# Is it normally distributed?
hist(aug$city_development_index, breaks = 20, freq = F)
curve(dnorm(x, mean(aug$city_development_index), sd(aug$city_development_index)), add = T)
# Shapiro test only accepts sample sizes between 3 and 5000
shapiro.test(aug$city_development_index[sample(1:nrow(aug), replace=FALSE, size=5000)])

# Detect outlier
out <- calculate_outliers(aug$city_development_index)
# Print number of extreme outliers
out$num_extreme
# Print number of mild outliers
out$num_mild

# Print the Barplot of the discreatized version of the variable
barplot(table(aug$f.city_development_index))
```
**Variable 4: gender**

The variable gender is a categorical variable with 4 levels that indicates the gender of employee. There are no explicit NA's, but the empty string "" appears to encode missing values. This variable is highly imbalanced, the number of male employees is 13221, whereas the numbers of female and other employees are 1238 and 191 respectively. By the way, the number of missing values defined by the empty string "" is 4508.

```{r}
# Summary of the column
sort(summary(factor(aug$gender)))
# Any missing values?
sum(is.na(aug$gender))
# Number of unique values
length(unique(aug$gender))
# Plot the variable
plot(aug$gender)

```
**Variable 5: relevent_experience**

The variable relevent_experience is a categorical variable with 2 levels that indicates the relevant job experience of employee. This variable has no missing values. There are 13792 employees with relevant experience and 5366 employees with no relevant experience. The number of the former is about twice and a bit more than the latter. The variable is imbalanced.

```{r}
# Summary of the column
sort(summary(aug$relevent_experience))
# Any missing values?
sum(is.na(aug$relevent_experience))
# Number of unique values
length(unique(aug$relevent_experience))
# Plot the variable
plot(aug$relevent_experience)

```
**Variable 6: enrolled_university**

The variable enrolled_university is a categorical variable with 4 levels that indicates the university enrollment status of the employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The number of missing values here is 386, the number of employees enrolled in a full time course is 3757, while those who did not enroll and enrolled in a part time course are 13817 and 1198 respectively. The variable is imbalanced, most of the employees did not enroll in any course.

```{r}
# Summary of the column
sort(summary(aug$enrolled_university))
# Any missing values?
sum(is.na(aug$enrolled_university))
# Number of unique values
length(unique(aug$enrolled_university))
# Plot the variable
plot(aug$enrolled_university)

```
**Variable 7: education_level**

The variable education_level is a categorical variable with 6 levels that indicates the highest education level of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Graduate 11598; High School 1017; Masters 4361; Phd 414; Primary School 308 and "" 460. The variable is imbalanced, most of the employees are graduates.

```{r}
# Summary of the column
sort(summary(aug$education_level))
# Any missing values?
sum(is.na(aug$education_level))
# Number of unique values
length(unique(aug$education_level))
# Plot the variable
plot(aug$education_level)
```
**Variable 8: major_discipline**

The variable major_discipline is a categorical variable with 7 levels that indicates the major discipline of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Arts 253;  Business Degree 327; Humanities 669; No Major 223; Other 381; STEM 14492 and "" 2813. We identified a cross-variable inconsistency. Employees with ‘Primary school’ or ‘High school’ education exceed those labeled ‘No major’ + ‘Other’ in major_discipline. This gap likely reflects the large volume of missing values in major_discipline. This variable is also imbalanced, most of the employees's major discipline is STEM.

```{r}
# Summary of the column
sort(summary(aug$major_discipline))
# Any missing values?
sum(is.na(aug$major_discipline))
# Number of unique values
length(unique(aug$major_discipline))

# Cross-variable inconsistency detected
ll1 <- which(aug$education_level == "High School" | aug$education_level == "Primary School"); length(ll1)
ll2 <- which(aug$major_discipline == "No Major" | aug$major_discipline == "Other"); length(ll2)
length(ll1) == length(ll2)

# Plot the variable
plot(aug$major_discipline)
```
**Variable 9: experience**

The variable major_discipline is a categorical variable with 23 levels that indicates the job experience of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: <1 522;  1 549; 2 1127; 3 1354; 4 1403; 5 1430; 6 1216; 7 1028; 8 802; 9 980; 10 985; 11 664; 12 494; 13 399; 14 586; 15 686; 16 508; 17 342; 18 280; 19 304; 20 148; >20 3286 and "" 65. This variable is also imbalanced, most of the employees's job experience is greater than 20 years.

```{r}
# Summary of the column
sort(summary(aug$experience))
# Any missing values?
sum(is.na(aug$experience))
# Number of unique values
length(unique(aug$experience))

# Plot the variable
plot(factor(aug$experience))
```
**Variable 10: company_size**

The variable company_size is a categorical variable with 9 levels that indicates the size of company that employee belong to. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: <10 1308;  10/49 1471;  50-99 3083; 100-500 2571; 500-999 877;  1000-4999 1328; 5000-9999 563; 10000+ 2019 and "" 5938 The data imbalance in this variable is less prevailing compared with other variables, since here the group with the highest number of observations is the empty string "".

```{r}
# Summary of the column
sort(summary(aug$company_size))
# Any missing values?
sum(is.na(aug$company_size))
# Number of unique values
length(unique(aug$company_size))

# Plot the variable
plot(factor(aug$company_size))
```
**Variable 11: company_type**

The variable company_type is a categorical variable with 7 levels that indicates the type of company that employee belong to. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Other 121;  Public sector 955; Early Stage Startup 603; Funded Startup 1001; NGO 521; Pvt Ltd 9817 and "" 6140. The variable is highly imbalanced, not only do most of the employees belong to Pvt Ltd, but also there are a lot of missing data.

```{r}
# Summary of the column
sort(summary(aug$company_type))
# Any missing values?
sum(is.na(aug$company_type))
# Number of unique values
length(unique(aug$company_type))

# Plot the variable
plot(factor(aug$company_type))
```
**Variable 12: last_new_job**

The variable last_new_job is a categorical variable with 7 levels that indicates difference in years between previous job and current job. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: never 2452;  1 8040;  2 2900; 3 1024; 4 1029;  >4 3290 and "" 423. The prevailing value here is 1.

```{r}
# Summary of the column
sort(summary(aug$last_new_job))
# Any missing values?
sum(is.na(aug$last_new_job))
# Number of unique values
length(unique(aug$last_new_job))

# Plot the variable
plot(aug$last_new_job)
```

**Variable 13: training_hours**

The variable price is a numerical variable indicating the training hours of employee. It ranges from 1 to 336. This variable has no missing values and there are 241 unique values. Based on histogram and Shapiro test we can consider it isn't normally distributed. Moreover, this variable have 275 extreme outlier and 984 mild outliers. In addition, looking at the barplot created using the discretized variable, we can say that all the groups have similar number of observations.

```{r}
# Summary of the column
summary(aug$training_hours)
# Any missing values?
sum(is.na(aug$training_hours))
# Number of unique values
length(unique(aug$training_hours))

# Is it normally distributed?
hist(aug$training_hours, breaks = 20, freq = F)
curve(dnorm(x, mean(aug$training_hours), sd(aug$training_hours)), add = T)
# Shapiro test only accepts sample sizes between 3 and 5000
shapiro.test(aug$training_hours[sample(1:nrow(aug), replace=FALSE, size=5000)])

# Detect outlier
out <- calculate_outliers(aug$training_hours)

# Print number of extreme outliers
out$num_extreme
# Print number of mild outliers
out$num_mild

# Print the Barplot of the discreatized version of the variable
barplot(table(aug$f.training_hours))
```

**Variable 14: target**

The variable target is a categorical (binary) variable with 2 levels that indicates job-change intent. There are no missing values and the counts here are: YES-Job-Change 4777 and NO-Job-Change 14381. This variable is highly imbalanced, there are around 10000 employees having no job-change intent more than those who have.

```{r}
# Summary of the column
sort(summary(aug$target))
# Any missing values?
sum(is.na(aug$target))
# Number of unique values
length(unique(aug$target))

# Plot the variable
plot(aug$target)
```
### 1.2. Data Quality Report

#### 1.2.1. Per Variable

Up to this points, all variables in the dataset have been explored. In the following cell, we will analyze the number of missing values, the number of inconsistencies and the number of outliers. Then, we will sum these values up and sort the variables in descending order.

Regarding the inconsistencies in the dataset, we assume that employees that did not enroll in any courses can neither be a graduate, a masters nor a phD as their highest level of education. Moreover, if the highest education level is Primary School, then they should have No Major in major_discipline variable.

```{r}
# Create an empty dataFrame
aug_var <- data.frame(missing_values = rep(NA, ncol(aug)-2))
aug_varnames <- setdiff(colnames(aug), c("f.city_development_index", "f.training_hours"))
rownames(aug_var) <- aug_varnames

# Count missing values
mss_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    sum(x == "")
  } else {
    0
  }
})
# Store missing values in the variables
aug_var$missing_values <- mss_list

# Count inconsistencies
err_list <- rep(0, length(aug_varnames))
err_list[which(aug_varnames=="enrolled_university" | aug_varnames=="education_level")] <-
        length(which(aug$enrolled_university == "no_enrollment" &
                    (aug$education_level == "Graduate" |
                     aug$education_level == "Masters" |
                     aug$education_level == "Phd")))
err_list[which(aug_varnames=="education_level")] <- err_list[which(aug_varnames=="education_level")] +
        length(which(aug$education_level == "Primary School" &
                     aug$major_discipline == "No Major"))
err_list[which(aug_varnames=="major_discipline")] <-
        length(which(aug$education_level == "Primary School" &
                     aug$major_discipline == "No Major"))
# Store erros in the variables
aug_var$errors <- err_list

# Count number of outliers
severe_out_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    return (0)
  } else {
    res.out <- calculate_outliers(x, plot_yes = FALSE)
    return (res.out$num_extreme)
  }
})
mild_out_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    return (0)
  } else {
    res.out <- calculate_outliers(x, plot_yes = FALSE)
    return (res.out$num_mild)
  }
})
# Store missing values in the variables
aug_var$severe_outliers <- severe_out_list
aug_var$mild_outliers <- mild_out_list
aug_var$total_outliers <- severe_out_list + mild_out_list

# Summing up columns
aug_var$sum_cols <- rowSums(aug_var)
# Sort the dataFrame in descending order
aug_var <- aug_var[order(aug_var$sum_cols, decreasing = TRUE), ]

# Print the dataFrame
aug_var
```

#### 1.2.2. Per Individual

Now, we will do something similiar with observations in the dataset. In the following cells, we will analyze the number of missing values, the number of inconsistencies and the number of outliers. Then we will identify the individuals considered as multivariant outliers.

```{r}
# Create an empty dataFrame
aug_ind <- data.frame(missing_values = rep(NA, nrow(aug)))

# Count missing values
mss_list <- apply(aug[row.names(aug_ind), ], MARGIN = 1, FUN = function(x) {
  sum(x == "")
})
# Store missing values in the variables
aug_ind$missing_values <- mss_list

# Count inconsistencies
# No inconsistencies found

# Count number of outliers
# Cannot compute outliers for individuals, since there are categorical values in each row

# Check preliminary multivariant outliers for the dataset
mout.res <- Moutlier(aug[,c(3,13)], plot = FALSE) # The missing data are not imputed yet
llmout <- which(mout.res$md > mout.res$cutoff & mout.res$rd > mout.res$cutoff) # ids of mout
# Store the information about multivariant outliers
aug_ind$mvout <- 0
aug_ind$mvout[llmout] <- 1

# Summing up columns
aug_ind$sum_cols <- rowSums(aug_ind)

# Print the dataFrame
aug_ind
```

Now, we want to study which variables are correlated with this final column of summed up missing values, errors and outliers. Since it is a numerical variable, we can use condes() from FactoMineR library to do the profiling. The numerical variables associated with sum_cols are training_hours and city_development_index, but their correlation with the variable is pretty low. On the other hand, the most correlated categorical variables with the variable sum_cols are those with a p-value equals 0 and a high number of R2. Just for mentioning, the top three most correlated categorical variables are company_size (0.628), company_type (0.62) and education_level (0.33).

```{r}
# Add the column to the original dataset
aug$sum_cols <- aug_ind$sum_cols

# Profiling using condes
res.con <- condes(aug[2:ncol(aug)], 16) # We skip the first column of ids

# Correlated numerical variables
res.con$quanti

# Correlated categorical variables
res.con$quali

# The effect of the categories is not interested here, since there are still NAs  
#res.con$category
```

In addition, we compute for every group of individuals the mean of missing values and then, rank the groups according to the computed mean. We observe that groups of empty string "" have the highest average number of missing values.

```{r}
# Group means of f.city_development_index
sort(tapply(aug$sum_cols, aug$f.city_development_index, mean))
print("===========================================")
# Group means of gender
sort(tapply(aug$sum_cols, aug$gender, mean))
print("===========================================")
# Group means of relevent_experience
sort(tapply(aug$sum_cols, aug$relevent_experience, mean))
print("===========================================")
# Group means of enrolled_university
sort(tapply(aug$sum_cols, aug$enrolled_university, mean))
print("===========================================")
# Group means of education_level
sort(tapply(aug$sum_cols, aug$education_level, mean))
print("===========================================")
# Group means of major_discipline
sort(tapply(aug$sum_cols, aug$major_discipline, mean))
print("===========================================")
# Group means of f.training_hours
sort(tapply(aug$sum_cols, aug$f.training_hours, mean))
```

```{r}
# We remove this last column in the dataset, since we will not be working with it from now on
aug$sum_cols <- NULL
```

### 1.3. Data Cleaning And Imputation

### 1.3.1. Numeric Variables

The only numerical variables we have in our dataset are city_development_index and training_hours. Since we only have two variables and the former one does not have severe outliers, we will remove all the mild outliers found using city_development_index and all the severe outliers found using training_hours.

```{r}
# Compute outliers
out_cdi <- calculate_outliers(aug$city_development_index, plot_yes = FALSE);
length(out_cdi$mild_indexes)
length(out_cdi$extreme_indexes)
out_th <- calculate_outliers(aug$training_hours, plot_yes = FALSE)
length(out_th$mild_indexes)
length(out_th$extreme_indexes)
```

```{r}
# Summary of the data before removing the outliers
summary(aug)
```

```{r}
# Indexes to remove
remove_indexes <- unique(c(out_cdi$mild_indexes, out_th$extreme_indexes))
length(remove_indexes)

# Summary of the outliers to remove
summary(aug[remove_indexes,])
```

```{r}
# Removing outliers
aug <- aug[-remove_indexes,]

# Summary of the cleaned dataset
summary(aug)

# Resetting row names
row.names(aug) <- NULL
```

Having printed the summaries of the outliers and the dataset without outliers, we can conclude that the outliers are essentially those observations with unusually low values of city development index and unusually high values of training hours.

### 1.3.2. Factors

Firstly, we need to assign NAs to the empty strings "" in all categorical variables. Then, we will apply imputeMCA() to impute categorical variables.

```{r}
# Assigning NAs
aug$gender[which(aug$gender == "")] <- NA
aug$enrolled_university[which(aug$enrolled_university == "")] <- NA
aug$education_level[which(aug$education_level == "")] <- NA
aug$major_discipline[which(aug$major_discipline == "")] <- NA
aug$experience[which(aug$experience == "")] <- NA
aug$company_size[which(aug$company_size == "")] <- NA
aug$company_type[which(aug$company_type == "")] <- NA
aug$last_new_job[which(aug$last_new_job == "")] <- NA

aug$gender <- factor(aug$gender)
aug$enrolled_university <- factor(aug$enrolled_university)
aug$education_level <- factor(aug$education_level)
aug$major_discipline <- factor(aug$major_discipline)
aug$experience <- factor(aug$experience)
aug$company_size <- factor(aug$company_size)
aug$company_type <- factor(aug$company_type)
aug$last_new_job <- factor(aug$last_new_job)
```

Note that the variables gender (4433), company_size (5852), and company_type (6051) have a lot of missing values in these columns. Although the variable major_discipline (2773) also has quite a lot, compared to the others its number is relatively low. An amount of around 5000 rows is too much for imputation, since it composes one quarter of the whole dataset, and the imputation results might not be accurate because they are assigned to the majority class, so we are going to assign an “Unknown” level to these three variables.

```{r}
# Imputation with imputePCA (~ 25 seconds)
res.mca <- imputeMCA(aug[, c(2,4:12, 14)])

# Missing data vs Imputed data
summary(aug[,c(2,4:12, 14)])
print("====================================")
summary(res.mca$completeObs)
```

```{r}
# Function to replace NAs with "Unknown"
add_unknown <- function(x) {
  x <- factor(x)
  levels(x) <- c(levels(x), "Unknown") # add new level
  x[is.na(x)] <- "Unknown" # replace NA
  return(x)
}

# Assigning "Unknown" levels
aug$gender <- add_unknown(aug$gender)
aug$company_size <- add_unknown(aug$company_size)
aug$company_type <- add_unknown(aug$company_type)

# Imputation with imputePCA
new.res.mca <- imputeMCA(aug[, c(2,5:9,12,14)]) # We exclude these three variables

# Missing data vs Imputed data
summary(aug[, c(2,5:9,12,14)])
print("====================================")
summary(new.res.mca$completeObs)

# Retain the imputed values
aug[, c(2,5:9,12,14)] <- new.res.mca$completeObs
```

Having compared the imputed and unimputed categorical variables, we notice that in variables with a large number of NAs (thousands of empty strings), the imputation process tends to assign most of these NAs to the majority class (we are mainly talking about major_discipline). This could be an error, or it might simply reflect the fact that the unnamed category genuinely belongs to the majority class. Either way, we will keep the imputed values and now, we will proceed with the rest of the assignment.

### 1.3.3. Removing Multivariate Outliers

Here, we use Moutlier() again to find multivariate outliers in the dataset and remove them from the dataset, since they can possibly decrease the performance of our final model.

```{r}
# Check multivariant outliers for the dataset
mout.res <- Moutlier(aug[,c(3,13)], plot = FALSE)
llmout <- which(mout.res$md > mout.res$cutoff & mout.res$rd > mout.res$cutoff) # ids of mout
length(llmout) # 687 multivariate outliers are found

# Comparing the values of the multivariate outliers
summary(aug[llmout,]) # Outliers
print("====================================")
summary(aug[-llmout,]) # No outliers

# Remove the multivariate outliers
aug <- aug[-llmout,]

# Resetting row names
row.names(aug) <- NULL
```

The multivariate outliers we found form a much smaller subset compared to the full dataset, which is expected. The categorical variables keep roughly the same proportions, so the main differences show up in the two numerical variables. The outliers seem to have lower values for city development index, since removing them increases the minimum from 0.479 to 0.512 and also raises the median to 0.91. For training hours, the outliers appear to take very high values, as the first quartile in the outlier set is already 202, almost at the upper end of the cleaned dataset. Overall, the outliers identified are those with unusually low city development index and unusually high training hours, similar to what we observed in the section 1.3.1.

### 1.3.4. Final Data Cleaning Process

In this section we will prepare the data a little bit more in order to facilitate the later analysis. In the Univariate Descriptive Analysis section we found different problems of each variable, so here we are going to either apply different cleaning functions to the columns or just remove them.

```{r}
# We remove the column enrollee_id which do no provide any useful information
aug <- aug[,setdiff(names(aug),"enrollee_id")]

# Check remaining columns
names(aug)
```

Then, we should check if there are any duplicated observations in the dataset. Since the column enrollee_id is removed, now we can check that. We cannot do it previously because the column enrollee_id has the same number of unique values as the number of rows in the dataset. For that purpose, we can use the function duplicated() which returns all rows that appear at least twice in the dataset. It seems that there are 52 duplicated observations in our current dataset.

```{r}
# Find number of duplicated rows
sum(duplicated(aug)) # 52 (If we did not assign "Unknown" levels, it would be 98)
# Find which observations are duplicated
duples <- which(duplicated(aug)) # One of the duplicated rows is in the position 240

# The function shows ids of rows that have already appeared before

# Show the duplicated observations of the observation 240
find_previous_duplicate <- function(df, row_id) {
  target <- df[row_id,]
  # Compare each row above row_id
  above_rows <- 1:(row_id - 1)
  for (i in above_rows) {
    row_i <- df[i,]
    # Compare element-wise, treating NA as equal
    equal <- mapply(function(a,b) a == b, row_i, target)
    if (all(equal)) return(i) # First match found
  }
  return(NA) # No previous match
}
previous_row_index <- find_previous_duplicate(aug, duples[1]) # Obs 240
aug[c(duples[1], previous_row_index),]

# Remove duplicated rows
aug <- aug[-duples,]

# Resetting row names
row.names(aug) <- NULL
```

The variables experience and company_size have too many levels, which will decrease the performance of the later modelling process. Therefore, we are going to group those levels into 4 groups at most (1 extra position for "Unknown" levels).

```{r}
# Summary of experience
summary(aug$experience)

# Grouping
aug$f.experience[aug$experience %in% c("<1", "1", "2", "3", "4", "5")] <- "[<1,5]"
aug$f.experience[aug$experience %in% c("6", "7", "8", "9", "10")] <- "[6,10]"
aug$f.experience[aug$experience %in% c("11", "12", "13", "14", "15")] <- "[11,15]"
aug$f.experience[aug$experience %in% c("16", "17", "18", "19", "20", ">20")] <- "[16,>20]"

# To factor
aug$f.experience <- factor(aug$f.experience)

# Print
aug$f.experience
```

```{r}
# Summary of company_size
summary(aug$company_size)

# Small: <10, 10/49, 50-99
aug$f.company_size[aug$company_size %in% c("<10", "10/49", "50-99")] <- "Small"
# Medium: 100-500, 500-999
aug$f.company_size[aug$company_size %in% c("100-500", "500-999")] <- "Medium"
# Large: 1000-4999, 5000-9999
aug$f.company_size[aug$company_size %in% c("1000-4999", "5000-9999")] <- "Large"
# Very Large: 10000+
aug$f.company_size[aug$company_size == "10000+"] <- "Very-Large"
# Unknown
aug$f.company_size[aug$company_size == "Unknown"] <- "Unknown"

# To factor
aug$f.company_size <- factor(aug$f.company_size)

# Print
aug$f.company_size
```

The variable city has 122 levels, which reduces the interpretability of subsequent models. We have two options: either remove it entirely or group it into a smaller number of categories. The most frequent cities are 103, 21, 16, 114, and 160, in that order. Since the remaining cities appear much less frequently, we can group them together with city 160. This way, we reduce the variable to just five meaningful categories, improving interpretability while retaining the most important information.

```{r}
# Number of city levels
lencity <- length(levels(aug$city))

# Which are the most frequent levels
sort(table(aug$city))[(lencity-4):lencity]

# Reducing the number of categories of city
aug$f.city <- 0
aug$f.city[which(aug$city == "city_103")] <- 4
aug$f.city[which(aug$city == "city_21")] <- 3
aug$f.city[which(aug$city == "city_16")] <- 2
aug$f.city[which(aug$city == "city_114")] <- 1

# From numeric to factors
aug$f.city <- factor(aug$f.city,labels=c("city_0","city_1","city_2","city_3","city_4"))

# Print the Barplot of the f.city
barplot(table(aug$f.city))
```

## 2. Profiling and Feature Selection

Our target variable is (binary) categorical, so we will use catdes() from FactoMineR library to do the profiling. The most correlated categorical variables with the target variable target are f.city_development_index, f.city, f.company_size, f.experience, enrolled_university, relevent_experience and education_level (p.value < 1e-30). On the other hand, the most correlated numerical variables (the only ones we have) are city_development_index (Eta2 = 0.11) and training_hours (Eta2 = 0.00044).

According to the results in \$quanti, employees who have the intent of changing the job have the significant mean lower than the grand mean in the variables training_hours and city_development_index. Next, let's focus on the results given by \$category. The employees who have the intent of changing the job are from the city 3 (33% of the group) and have low city_development_index (49% of the group) and their current company's size and type are unknown (50% of the group for each variable). Their total experience in years is between less than a year and 5 years (46% of the group), so they do not have relevant experience (38% of the group), they were enrolled in a full time course at a university (30% of the group) and their education level is bachelor's (70% of the group). Based on this information, we can conclude that individuals seeking to change jobs are likely recent graduates with limited experience in the field.

```{r}
# Applying catdes on target
res.cat <- catdes(aug[,c(2:7,10:ncol(aug))], num.var=10)
# We exclude city, experience and company_size

# Most significant categorical variables
res.cat$test.chi2

# Most significant numerical variables (the only two)
res.cat$quanti.var

# Description by the numerical variables
res.cat$quanti

# Description by the categorical variables
res.cat$category
```

In the posterior modelling, we will focus on the most significant variables identified during the profiling of the binary target. Both numerical variables and the discretized version of city_development_index are highly significant. However, since we only have two numerical variables in total, it makes more sense to keep city_development_index as a continuous variable and use it together with training_hours.

On the other hand, all remaining categorical variables also appear to be strongly significant. The variables city, experience and company_size contain too many levels which will unnecessary increase the interpretability of the models, so we will use f.city, f.experience and f.company_size instead. 

## 3. Modeling Using Numeric Variables

Binary outcome modelling is widely used in real situations, such as deciding whether to grant a loan or predicting if a player will get injured in the next match. To handle this type of problem, we rely on logistic regression, where the key component is the link function, which transforms the linear predictor from the range $(-\infty, \infty)$ into a probability between 0 and 1. In our case, we will start by using the two relevant numerical variables in the dataset to build a binary model using glm().

We have two possible link functions: **Logit** and **Probit**

#### **(1) Logit Link Function**

The logit link transforms a linear predictor from the real line $(-\infty, \infty)$  
into a probability in the interval $[0, 1]$:

\[
\text{logit}(\pi) = \ln\left(\frac{\pi}{1 - \pi}\right)
\]

The inverse link (**logistic function**) gives the predicted probability:

\[
g^{-1}(X) = \pi = \frac{\exp(X)}{1 + \exp(X)}
\]

Using a threshold \(s\), we convert probabilities into class predictions:

\[
\pi > s \Rightarrow \hat{y} = 1
\]

\[
\pi \le s \Rightarrow \hat{y} = 0
\]

---

#### **(2) Probit Link Function**

The probit link is based on the cumulative distribution function (CDF) of the standard normal distribution. It maps a linear predictor to a probability using:

\[
\pi = \Phi(X)
\]

In R:

- `pnorm()` computes the standard normal CDF, returning the probability that a normal variable is less than or equal to a given value:

```{r}
pnorm(-2.12, lower.tail = TRUE)
```

- `qnorm()` performs the inverse operation, returning the quantile corresponding to a given probability:

```{r}
qnorm(0.01700302)
```

In the next step, we will split the data into training and test sets. The test set will remain unseen during model building, so all model fitting and initial analysis will be performed primarily on the training set, while the test set will be reserved for evaluating model performance.

```{r}
# Set a seed to reproduce outcomes
set.seed(07150523)

# Split dataset
sets <- sample(1:nrow(aug),size = round(0.8*nrow(aug), dig=0), replace = FALSE)
aug_train <- aug[sets,]
aug_test <- aug[-sets,]
```

Before using the numerical predictors, we will analyze the NULL model derived from the target variable target. The target variable has two levels, NO-Job-Change and YES-Job-Change, with proportions 0.7517 and 0.248 respectively in the dataset. By computing the log of the odds of the target class (YES-Job-Change), we obtain the intercept of the NULL model. The odds of success is 0.33, which means we have around 0.33% of chance to find a success than a failure, the negative value of intercept (-1.107643) means it is more likely to find failure classes (NO-Job-Change).

```{r}
# NULL model, by default we are using the logit link function
m0 <- glm(target ~ 1, family=binomial, data = aug_train)

# Proportions of NoJobChange vs YesJobChange
ptt<-prop.table(table(aug_train$target))
ptt # Estimation of the probability of having a job change

# Odds of the positive class: How probable is it to find a success than a failure
odds <- ptt[2]/(1-ptt[2])
odds

# log(Odds)
log(odds) # This gives us the intercept of the NULL model
# Positive: more likely to find successful events
# Negative: more likely to find unsuccessful events

# The intercept or log(Odds) can also be computed using qlogis (when we use logit)
qlogis(ptt)[2] # Intercept of the positive class

# From the intercept we can get the probability applying the inverse link function
lp <- log(odds) # qlogis(ptt)[2]
exp(lp) / (1 + exp(lp)) # The estimation of the probability of having a job change
# This can also be done using the following function
plogis(log(odds))

# The intercept estimate is the log of odds of the NULL model
summary(m0)
```

In case of using probit link function, the intercept we get is different. In order to get this value here, we need to apply qnorm() on the estimated probability of the positive class and then apply pnorm on the intercept to get the same probability we get in the previous cell. This shows that even though the intercept and predictor values change from one link function to another, the resulting predicted probabilities remain the same. For the rest of the exercises, we will focus on only one link function; in our case, we will use the logit function.

```{r}
# NULL mode (probit)
m0p <- glm(target ~ 1, family=binomial(link="probit"), data = aug_train)
summary(m0p)

# The intercept or log(Odds)
qnorm(ptt)[2]
# Get the probability from the intercept
pnorm(coef(m0p)[1]) # pnorm(qnorm(ptt)[2])
```

Next, we add the numerical predictors to the model. When we check the summary of model m1, we see that the estimate for training_hours is very small. This by itself is not an issue; what matters is that the p-value for this predictor is very high (above 0.01 or 0.05), indicating that it is not statistically significant. Because of this lack of significance, we will discard training_hours from the model.

The estimate of the predictor city_development_index is -6.02, which is the new effect added to the model. This means that when the city development index increases by one unit, the log(odds) of YES-Job-Change decrease by about 6.02. In other words, higher city development index values are associated with a lower probability of job change according to the model. To support this idea, we computed the predicted probabilities of job change for a high value of city_development_index (0.949) and for a low value (0.512). The results show that when the index is high, the probability of job change is approximately 0.1262, whereas when the index is low, the probability increases to about 0.668.

```{r}
# Adding numeric predictors to the model
m1 <- glm(target ~ city_development_index + training_hours, family=binomial(link="logit"), data=aug_train)
summary(m1)

# New model after discarding training_hours
m1d <- glm(target ~ city_development_index, family=binomial(link="logit"), data=aug_train)
summary(m1d)

# ------------------------------------------------------------------------------

# Total estimation (log-odds) of job change when city_development_index is high
lp_max <- coef(m1d)[1] + 0.949 * coef(m1d)[2] # 0.949 = maximum value in the dataset
lp_max
# Get the probability from the linear predictor lp_max
exp(lp_max) / (1 + exp(lp_max)) # Probability of job change when the index is high

# Total estimation (log-odds) of job change when city_development_index is low
lp_min <- coef(m1d)[1] + 0.512 * coef(m1d)[2] # 0.512 = minimum value in the dataset
lp_min
# Get the probability from the linear predictor lp_min
exp(lp_min) / (1 + exp(lp_min)) # Probability of job change when the index is low
```

Next, we check if we should apply any transformation to city_development_index. For this purpose, we will use boxcox. The lambda we obtain is 2, which suggests that we should apply a square effect to the variable. After applying the square effect, we can see through the summary function that all coefficients in the model m1s are statistically significant according to the low p-value.

```{r}
# Applying boxcox
b <- boxcox(city_development_index ~ 1, data = aug_train)
lambda <- b$x[which.max(b$y)]; lambda

# Applying square effect
m1s <- glm(target ~ poly(city_development_index, 2), family=binomial(link="logit"), data=aug_train)
summary(m1s)
```

When comparing the residual plots for the quadratic and linear models, the quadratic model shows a better fit in the left plot: the residuals follow the curvature of the underlying relationship more closely. However, in the right plot the linear predictor displays an odd pattern, the pink line first dips and then rises into the positive Pearson residuals, which suggests the added squared term may be introducing some distortion.

On the other hand, the marginal model plots show that the model with the quadratic term captures the curvature of the data more effectively. Its fitted line adapts well to the underlying nonlinear pattern, whereas the linear model does not reproduce this shape. Nonetheless, the linear fit still follows the general direction and overall trend of the data, even if it misses the curvature.

```{r}
# The models we get from the previous cell
m1 <- glm(target ~ poly(city_development_index, 2), family=binomial, data = aug_train)
# NULL model
m0 <- glm(target ~ 1, family=binomial, data = aug_train)
# The model without square effect on city_development_index
m1d <- glm(target ~ city_development_index, family=binomial, data=aug_train)
```

```{r}
# Which variables are important for modelling
Anova(m1, test="LR")
Anova(m1d, test="LR")
# Both linear and quadratic effects of city_development_index are significant
```

```{r}
# ResidualPlots for m1 and m1d
residualPlots(m1)
residualPlots(m1d)
```

```{r}
# MarginalModelPlots for each model
marginalModelPlots(m1)
marginalModelPlots(m1d)
```

Based on the Analysis of Deviance using anova(), the model with the quadratic term provides a significantly better fit than the simple linear model. The likelihood ratio test shows a large difference in deviance (71.72) with an extremely small p-value (< 2.2e-16), indicating that the quadratic term explains variation in the response that the linear term alone cannot capture. Therefore, the quadratic model is preferred, as it better accounts for the nonlinear relationship between city_development_index and the probability of the target outcome.

```{r}
# Apply a Deviance test
anova(m1d, m1, test="Chisq")
```

Next, we explored adding a cubic term for city_development_index to see if it further improves the model. All coefficients remained statistically significant. We also tried higher-order polynomial transformations, such as a 4th-degree polynomial, and again all coefficients were significant. Although we could continue increasing the polynomial degree, we chose to stop to avoid overfitting the model. Additionally, the step() function confirmed that none of the coefficients were insignificant, supporting the inclusion of the selected polynomial terms.

The residual plots of the cubic and 4th-degree polynomial models are quite similar, but if we had to choose, the cubic model appears preferable. The marginal model plots of the cubic model show that it closely follows the pattern of the data, whereas the 4th-degree polynomial model aligns almost perfectly. While this may seem desirable, it raises the risk of overfitting. The difference in deviance between the two models is 25.524, indicating that the higher-order model provides only a modest improvement in fit compared to the cubic model. This conclusion is further supported by the AIC values, where the more complex model only reduces AIC by 23.53 units (from 3th-degree to 4th-degree). Therefore, we decide to retain the cubic model, as it balances a good fit with lower complexity and a reduced risk of overfitting.

```{r}
# Adding cubic effect
m1c3 <- glm(target ~ poly(city_development_index, 3), family=binomial, data = aug_train)
summary(m1c3)
step(m1c3)

# Use 4th degree polynomial
m1c4 <- glm(target ~ poly(city_development_index, 4), family=binomial, data = aug_train)
summary(m1c4)
step(m1c4)
```

```{r}
# ResidualPlots for m1c3 and m1c4
residualPlots(m1c3)
residualPlots(m1c4)
```

```{r}
# MarginalModelPlots for each model
marginalModelPlots(m1c3)
marginalModelPlots(m1c4)
```

```{r}
# Apply a Deviance test
anova(m1c3, m1c4, test="Chisq")
```

```{r}
AIC(m1c3,m1) # The m1c3 model decreases AIC by 50, compared with the square effect model
AIC(m1c4,m1) # The m1c4 model decreases AIC by 73.5
AIC(m1c3,m1c4) # The m1c4 model only provides a small decrease in AIC (23.53)
```

### 3.1. Residual Analysis: Unusual and Influent Data Filtering

In this section, we remove unusual and influential observations from the training set to refine the model. We begin by examining the marginal model plots and residual plots of the cubic model, which already shows a good overall fit and follows the nonlinear pattern in the data reasonably well. Nonetheless, some observations exert disproportionately large influence, so we apply influencePlot, hat-values, and Cook’s distance to identify such cases. After removing the most extreme influential observations, the marginal model plots show that the refitted cubic model aligns slightly better with the data. This improvement is also reflected in a decrease of about 45.43 AIC units. When comparing Cook’s distances before and after cleaning, the original model contained points exceeding 0.002, whereas the cleaned model removes most of them. Since the remaining Cook’s distances, although some reach values around 0.002, are still very small relative to the sample size (~15,000 observations), these levels of influence are not problematic for a logistic regression model. Additionally, the coefficients of the cubic terms change by only about one unit, suggesting that the model has stabilized. For these reasons, we do not remove any further observations.

```{r}
# Fit cubic polynomial model
m1c3 <- glm(target ~ poly(city_development_index, 3), family=binomial, data = aug_train)

# ----------------------------
# Marginal model plots
#    - Shows the partial relationship between predictor and response
#    - id: label the top 5 most influential points according to Cook's distance
# ----------------------------
marginalModelPlots(m1c3,id=list(labels=row.names(aug_train),method=abs(cooks.distance(m1c3)), n=5))

# ----------------------------
# Residual diagnostics
#    - residualPlots: component + partial residual plots
#    - layout: arrange plots in 3 rows x 2 cols
# ----------------------------
residualPlots(m1c3, layout = c(3, 2))
```

```{r}
# ----------------------------
# Influence diagnostics (influencePlot)
#    - Identifies observations with large leverage and influence
#    - Returns row names of augmented training data
# ----------------------------
llnrem <- influencePlot(m1c3, id = list(n = 5)) # plots and returns the identified points
llnrem # shows which rows were flagged
attributes(llnrem) # row names
high_lever_influ_idx <- which(rownames(aug_train) %in% rownames(llnrem)) # row numbers
high_lever_influ_idx

# ----------------------------
# Apriori influential data analysis using hat-values and Cook's distance
#    - Boxplot of hat-values: identifies high-leverage points
#    - Add a reference line at 4*p/n
# ----------------------------
# hat-values boxplot: returns indices of labelled points (row numbers in aug_train)
high_leverage_idx <- Boxplot(hatvalues(m1c3), id.n = 10)
abline(h = 4 * length(coef(m1c3)) / nrow(aug_train), col = "red") # heuristic cutoff
hatvalues(m1c3)[high_leverage_idx]

# ----------------------------
# Cook's distance: detect influential observations
#    - Boxplot helps visualize and label the top Cook's distances
# ----------------------------
high_cooks_idx <- Boxplot(cooks.distance(m1c3), id.n = 10)
# Show which observations (rows) are influential
cooks.distance(m1c3)[high_cooks_idx]
aug_train[high_cooks_idx, ]
```

```{r}
# ----------------------------
# Create a cleaned training set removing the influent data
# ----------------------------
aug_train_clean <- aug_train[-unique(c(high_lever_influ_idx, high_leverage_idx, high_cooks_idx)), ]

# Fit the cubic polynomial model using the cleaned data
m1c32 <- glm(target ~ poly(city_development_index, 3), family=binomial, data = aug_train_clean)
```

```{r}
# ResidualPlots for m1c3 and m1c32
residualPlots(m1c3)
residualPlots(m1c32)
```

```{r}
# MarginalModelPlots for each model
marginalModelPlots(m1c3)
marginalModelPlots(m1c32)
```

```{r}
AIC(m1c3,m1c32) # The m1c32 model decreases AIC by 45 units, compared to m1c3
```

```{r}
# Check Cook’s distance before removal
plot(cooks.distance(m1c3))
abline(h = 4/nrow(aug_train), col="red")
```

```{r}
# Check Cook’s distance after removal
plot(cooks.distance(m1c32))
abline(h = 4/nrow(aug_train_clean), col="red")
```

```{r}
summary(m1c3)
```

```{r}
summary(m1c32)
```

## 4. Adding factor main effects to the best model

```{r}
m2 <- glm(target ~ poly(city_development_index, 3), family=binomial, data = aug_train_clean)
```












-| severe and mild outliers should not be set to NA (remove the severe ones only)
-| multivariate outliers should be removed (after analyzing the multivariate outliers)
-| check consistency after imputation / validation of the imputation
-| we need to show condes() catdes() outspust (all outputs)
-| we do not use factors with too many levels
-> eliminate severe multicollinearity (GVIFˆ (1/(2∗Df))>10) to the full
-> eliminate residual outliers to the full
-> non-normally distributed residuals in qq-plot
-> show cook's distance for influent data checking
-> R2 goodness of fit should be discussed in the final model
-> Residual analysis shows divergences to linear model hypothesis (fix this situation)
-> Influent data cleaning process should be shown


---> We do the splits after analyzing the whole/original dataset (although in the real world scenarios we can only use the train set to analyze the distribution of the data)

Then, we split it into train and test sets. The test set will have 15% of the original dataset (2873), while the train set will have the remaining 85% (16285).




