---
title: "Lab_Assignment_2"
author: "Zheshuo and Geming"
date: "2025-11-10"
output: pdf_document
---


# ASSIGNMENT 2: HR Analytics - Job Change of Data Scientists 

## 0. Before Starting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Clear the plots
rm(list=ls()) # Clear the environment
set.seed(07150523) # Set a seed to reproduce outcomes
```

We start by loading essential libraries and importing the data.

```{r}
# Installing dependencies
#install.packages("lubridate")
```

```{r include = FALSE}
# Loading libraries
library(moments)
library(car)
library(chemometrics)
library(mice)
library(missMDA)
library(FactoMineR)
library(corrplot)
library(VIM)
library(dplyr)
library(ggplot2)
library(AER)
library(MASS)
library(dplyr)
library(lubridate)
library(stringr)
library(readr)
library(tidyr)
library(tidyselect)
library(effects)
```

```{r}
# Importing the main data of the assignment
aug <- read.csv("aug_train.csv")
```

The dataset contains 19158 observations about different employees. Each row has 14 attributes/variables containing information related to demographics, education and experience of the employee.

```{r}
# Show the name of columns
names(aug)
```

```{r}
# Check the structure of the dataset
str(aug)
```

```{r}
# Here we convert the categorical variables into factors
aug$city <- factor(aug$city)
aug$gender <- factor(aug$gender)
aug$relevent_experience <- factor(aug$relevent_experience)
aug$enrolled_university <- factor(aug$enrolled_university)
aug$education_level <- factor(aug$education_level)
aug$major_discipline <- factor(aug$major_discipline)
aug$experience <- factor(aug$experience)
aug$company_size <- factor(aug$company_size)
aug$company_type <- factor(aug$company_type)
aug$last_new_job <- factor(aug$last_new_job)
```


```{r}
# Use summary function to get a fast understanding of the dataset
summary(aug)
```

```{r}
# Print the first 10 observations in the data
head(aug, 10)
```


## 1. Data Preparation

### 1.1. Univariate Descriptive Analysis

First of all, we can create a discretization version of variable for the the numerical variables city_development_index, training_hours and the target binary variable target. This can be done using the following cell. Moreover, the dataset has no apparent missing values (NAs), but we cannot discard the posibility of having them (missing values) as outliers in the dataset.

```{r}
# Creating the polytomic factors by city_development_index and training_hours quantiles
get_polytomic_factors <- function(x, variable, labels){
  # Compute quartile cut points
  qs <- quantile(variable, probs = c(0, 0.25, 0.5, 0.75, 1))
  
  # Derive a name for the new column (e.g. "f.city" if variable was aug$city)
  var_name <- deparse(substitute(variable))
  colname <- paste0("f.", substring(var_name, 5))
  
  # Create the factor column
  x[[colname]] <- cut(variable,
                      breaks = qs,
                      include.lowest = TRUE,
                      labels = labels)
  return(x)
}

aug <- get_polytomic_factors(aug, aug$city_development_index, labels = c("Low",
                                                                  "Med-Low",
                                                                  "Med-High",
                                                                  "High"))
aug <- get_polytomic_factors(aug, aug$training_hours, labels = c("Q1", "Q2", "Q3", "Q4"))
aug$target <- factor(aug$target, labels = c("NO-Job-Change", "YES-Job-Change"))

```

In the next cell we implement a function that calculates the univariate outliers of a given numeric variable in the dataframe object.

```{r}
# The function we need for EDA processes
# Use given Q1 & Q3 to compute indexes of outliers
calculate_outliers <- function(x, plot_yes = TRUE) {
  # Check for regular outliers
  col_name <- deparse(substitute(x))
  outliers <- boxplot.stats(x)$out

  # Compute quartiles and IQR
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1

  # Define thresholds
  lower_mild <- Q1 - 1.5 * IQR
  upper_mild <- Q3 + 1.5 * IQR
  lower_extreme <- Q1 - 3 * IQR
  upper_extreme <- Q3 + 3 * IQR

  # Identify mild and extreme outliers
  mild_idx <- which((x < lower_mild | x > upper_mild))
  extreme_idx <- which((x < lower_extreme | x > upper_extreme))
  
  # Count
  num_mild <- length(mild_idx)
  num_extreme <- length(extreme_idx)

  if (plot_yes) {
    Boxplot(x, main = "Boxplot", ylab = col_name)
    abline(h = c(lower_mild, upper_mild), col = "blue", lwd = 2)
    abline(h = c(lower_extreme, upper_extreme), col = "red", lwd = 2)
  }

  # Return the results as a list
  return(list(
    num_mild = num_mild,
    num_extreme = num_extreme,
    mild_indexes = mild_idx,
    extreme_indexes = extreme_idx
  ))
}

```

**Variable 1: enrollee_id**

Variable enrollee_id is a numerical variable with no missing values thus imputation is not needed. A quick inspection reveals that it serves as a unique identifier for each observation, since the number of unique values equals the total number of rows. Therefore, this variable does not contain meaningful numerical information for statistical analysis.

Normality and outlier tests were not performed, as the variable is an ID rather than a continuous numeric feature. Interestingly, the maximum value (33380) in the data does not match the number of columns (19158), which might mean that the dataset we have here is a subset of the original one. The variable enrollee_id ranges from 1 to 33380.

```{r}
# Summary of the column
summary(aug$enrollee_id)
# Any missing values?
sum(is.na(aug$enrollee_id))
# Number of unique values
unique_counts <- length(unique(aug$enrollee_id))
# Verify if all values are unique
unique_counts == nrow(aug)

```

**Variable 2: city**

The variable city is a categorical variable indicating the id of the city where the employees are located in. There are no missing data in the data and it has 123 unique values in this column. The most frequent city is city_103 with 4355 appearances and the lowest are city_127, city_133, city_167 and city_59 with 10 appearances.

```{r}
# Number of levels in the variable
length(levels(aug$city))
# Summary of the column
sort(summary(aug$city))
# Are there missing values?
sum(is.na(aug$city)) # No missing values

```

**Variable 3: city_development_index**

The variable city_development_index is a numerical variable indicating the development index of city. It ranges from 0.4480 to 0.9490. This variable has no missing values and there are 93 unique values. Based on histogram and Shapiro test we can consider it isn't normally distributed. Moreover, this variable doesn't have any extreme outlier but it has 17 mild outliers. In addition, looking at the barplot created using the discretized variable, we can say that the first two groups have similar number of observations. While the third group has the largest number of observations, the fourth group has the lowest.

```{r}
# Summary of the column
summary((aug$city_development_index))
# Any missing values?
sum(is.na(aug$city_development_index))
# Number of unique values
length(unique(aug$city_development_index))

# Is it normally distributed?
hist(aug$city_development_index, breaks = 20, freq = F)
curve(dnorm(x, mean(aug$city_development_index), sd(aug$city_development_index)), add = T)
# Shapiro test only accepts sample sizes between 3 and 5000
shapiro.test(aug$city_development_index[sample(1:nrow(aug), replace=FALSE, size=5000)])

# Detect outlier
out <- calculate_outliers(aug$city_development_index)
# Print number of extreme outliers
out$num_extreme
# Print number of mild outliers
out$num_mild

# Print the Barplot of the discreatized version of the variable
barplot(table(aug$f.city_development_index))
```
**Variable 4: gender**

The variable gender is a categorical variable with 4 levels that indicates the gender of employee. There are no explicit NA's, but the empty string "" appears to encode missing values. This variable is highly imbalanced, the number of male employees is 13221, whereas the numbers of female and other employees are 1238 and 191 respectively. By the way, the number of missing values defined by the empty string "" is 4508.

```{r}
# Summary of the column
sort(summary(factor(aug$gender)))
# Any missing values?
sum(is.na(aug$gender))
# Number of unique values
length(unique(aug$gender))
# Plot the variable
plot(aug$gender)

```
**Variable 5: relevent_experience**

The variable relevent_experience is a categorical variable with 2 levels that indicates the relevant job experience of employee. This variable has no missing values. There are 13792 employees with relevant experience and 5366 employees with no relevant experience. The number of the former is about twice and a bit more than the latter. The variable is imbalanced.

```{r}
# Summary of the column
sort(summary(aug$relevent_experience))
# Any missing values?
sum(is.na(aug$relevent_experience))
# Number of unique values
length(unique(aug$relevent_experience))
# Plot the variable
plot(aug$relevent_experience)

```
**Variable 6: enrolled_university**

The variable enrolled_university is a categorical variable with 4 levels that indicates the university enrollment status of the employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The number of missing values here is 386, the number of employees enrolled in a full time course is 3757, while those who did not enroll and enrolled in a part time course are 13817 and 1198 respectively. The variable is imbalanced, most of the employees did not enroll in any course.

```{r}
# Summary of the column
sort(summary(aug$enrolled_university))
# Any missing values?
sum(is.na(aug$enrolled_university))
# Number of unique values
length(unique(aug$enrolled_university))
# Plot the variable
plot(aug$enrolled_university)

```
**Variable 7: education_level**

The variable education_level is a categorical variable with 6 levels that indicates the highest education level of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Graduate 11598; High School 1017; Masters 4361; Phd 414; Primary School 308 and "" 460. The variable is imbalanced, most of the employees are graduates.

```{r}
# Summary of the column
sort(summary(aug$education_level))
# Any missing values?
sum(is.na(aug$education_level))
# Number of unique values
length(unique(aug$education_level))
# Plot the variable
plot(aug$education_level)
```
**Variable 8: major_discipline**

The variable major_discipline is a categorical variable with 7 levels that indicates the major discipline of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Arts 253;  Business Degree 327; Humanities 669; No Major 223; Other 381; STEM 14492 and "" 2813. We identified a cross-variable inconsistency. Employees with ‘Primary school’ or ‘High school’ education exceed those labeled ‘No major’ + ‘Other’ in major_discipline. This gap likely reflects the large volume of missing values in major_discipline. This variable is also imbalanced, most of the employees's major discipline is STEM.

```{r}
# Summary of the column
sort(summary(aug$major_discipline))
# Any missing values?
sum(is.na(aug$major_discipline))
# Number of unique values
length(unique(aug$major_discipline))

# Cross-variable inconsistency detected
ll1 <- which(aug$education_level == "High School" | aug$education_level == "Primary School"); length(ll1)
ll2 <- which(aug$major_discipline == "No Major" | aug$major_discipline == "Other"); length(ll2)
length(ll1) == length(ll2)

# Plot the variable
plot(aug$major_discipline)
```
**Variable 9: experience**

The variable major_discipline is a categorical variable with 23 levels that indicates the job experience of employee. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: <1 522;  1 549; 2 1127; 3 1354; 4 1403; 5 1430; 6 1216; 7 1028; 8 802; 9 980; 10 985; 11 664; 12 494; 13 399; 14 586; 15 686; 16 508; 17 342; 18 280; 19 304; 20 148; >20 3286 and "" 65. This variable is also imbalanced, most of the employees's job experience is greater than 20 years.

```{r}
# Summary of the column
sort(summary(aug$experience))
# Any missing values?
sum(is.na(aug$experience))
# Number of unique values
length(unique(aug$experience))

# Plot the variable
plot(factor(aug$experience))
```
**Variable 10: company_size**

The variable company_size is a categorical variable with 9 levels that indicates the size of company that employee belong to. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: <10 1308;  10/49 1471;  50-99 3083; 100-500 2571; 500-999 877;  1000-4999 1328; 5000-9999 563; 10000+ 2019 and "" 5938 The data imbalance in this variable is less prevailing compared with other variables, since here the group with the highest number of observations is the empty string "".

```{r}
# Summary of the column
sort(summary(aug$company_size))
# Any missing values?
sum(is.na(aug$company_size))
# Number of unique values
length(unique(aug$company_size))

# Plot the variable
plot(factor(aug$company_size))
```
**Variable 11: company_type**

The variable company_type is a categorical variable with 7 levels that indicates the type of company that employee belong to. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: Other 121;  Public sector 955; Early Stage Startup 603; Funded Startup 1001; NGO 521; Pvt Ltd 9817 and "" 6140. The variable is highly imbalanced, not only do most of the employees belong to Pvt Ltd, but also there are a lot of missing data.

```{r}
# Summary of the column
sort(summary(aug$company_type))
# Any missing values?
sum(is.na(aug$company_type))
# Number of unique values
length(unique(aug$company_type))

# Plot the variable
plot(factor(aug$company_type))
```
**Variable 12: last_new_job**

The variable last_new_job is a categorical variable with 7 levels that indicates difference in years between previous job and current job. There are no explicit Na's, but the empty string "" appears to encode missing values. The counts here are: never 2452;  1 8040;  2 2900; 3 1024; 4 1029;  >4 3290 and "" 423. The prevailing value here is 1.

```{r}
# Summary of the column
sort(summary(aug$last_new_job))
# Any missing values?
sum(is.na(aug$last_new_job))
# Number of unique values
length(unique(aug$last_new_job))

# Plot the variable
plot(aug$last_new_job)
```

**Variable 13: training_hours**

The variable price is a numerical variable indicating the training hours of employee. It ranges from 1 to 336. This variable has no missing values and there are 241 unique values. Based on histogram and Shapiro test we can consider it isn't normally distributed. Moreover, this variable have 275 extreme outlier and 984 mild outliers. In addition, looking at the barplot created using the discretized variable, we can say that all the groups have similar number of observations.

```{r}
# Summary of the column
summary(aug$training_hours)
# Any missing values?
sum(is.na(aug$training_hours))
# Number of unique values
length(unique(aug$training_hours))

# Is it normally distributed?
hist(aug$training_hours, breaks = 20, freq = F)
curve(dnorm(x, mean(aug$training_hours), sd(aug$training_hours)), add = T)
# Shapiro test only accepts sample sizes between 3 and 5000
shapiro.test(aug$training_hours[sample(1:nrow(aug), replace=FALSE, size=5000)])

# Detect outlier
out <- calculate_outliers(aug$training_hours)

# Print number of extreme outliers
out$num_extreme
# Print number of mild outliers
out$num_mild

# Print the Barplot of the discreatized version of the variable
barplot(table(aug$f.training_hours))
```

**Variable 14: target**

The variable target is a categorical (binary) variable with 2 levels that indicates job-change intent. There are no missing values and the counts here are: YES-Job-Change 4777 and NO-Job-Change 14381. This variable is highly imbalanced, there are around 10000 employees having no job-change intent more than those who have.

```{r}
# Summary of the column
sort(summary(aug$target))
# Any missing values?
sum(is.na(aug$target))
# Number of unique values
length(unique(aug$target))

# Plot the variable
plot(aug$target)
```
### 1.2. Data Quality Report

#### 1.2.1. Per Variable

Up to this points, all variables in the dataset have been explored. In the following cell, we will analyze the number of missing values, the number of inconsistencies and the number of outliers. Then, we will sum these values up and sort the variables in descending order.

Regarding the inconsistencies in the dataset, we assume that employees that did not enroll in any courses can neither be a graduate, a masters nor a phD as their highest level of education. Moreover, if the highest education level is Primary School, then they should have No Major in major_discipline variable.

```{r}
# Create an empty dataFrame
aug_var <- data.frame(missing_values = rep(NA, ncol(aug)-2))
aug_varnames <- setdiff(colnames(aug), c("f.city_development_index", "f.training_hours"))
rownames(aug_var) <- aug_varnames

# Count missing values
mss_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    sum(x == "")
  } else {
    0
  }
})
# Store missing values in the variables
aug_var$missing_values <- mss_list

# Count inconsistencies
err_list <- rep(0, length(aug_varnames))
err_list[which(aug_varnames=="enrolled_university" | aug_varnames=="education_level")] <-
        length(which(aug$enrolled_university == "no_enrollment" &
                    (aug$education_level == "Graduate" |
                     aug$education_level == "Masters" |
                     aug$education_level == "Phd")))
err_list[which(aug_varnames=="education_level")] <- err_list[which(aug_varnames=="education_level")] +
        length(which(aug$education_level == "Primary School" &
                     aug$major_discipline == "No Major"))
err_list[which(aug_varnames=="major_discipline")] <-
        length(which(aug$education_level == "Primary School" &
                     aug$major_discipline == "No Major"))
# Store erros in the variables
aug_var$errors <- err_list

# Count number of outliers
severe_out_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    return (0)
  } else {
    res.out <- calculate_outliers(x, plot_yes = FALSE)
    return (res.out$num_extreme)
  }
})
mild_out_list <- sapply(aug[,aug_varnames], function(x) {
  if (is.factor(x) || is.character(x)) {
    return (0)
  } else {
    res.out <- calculate_outliers(x, plot_yes = FALSE)
    return (res.out$num_mild)
  }
})
# Store missing values in the variables
aug_var$severe_outliers <- severe_out_list
aug_var$mild_outliers <- mild_out_list
aug_var$total_outliers <- severe_out_list + mild_out_list

# Summing up columns
aug_var$sum_cols <- rowSums(aug_var)
# Sort the dataFrame in descending order
aug_var <- aug_var[order(aug_var$sum_cols, decreasing = TRUE), ]

# Print the dataFrame
aug_var
```

#### 1.2.2. Per Individual

Now, we will do something similiar with observations in the dataset. In the following cells, we will analyze the number of missing values, the number of inconsistencies and the number of outliers. Then we will identify the individuals considered as multivariant outliers.

```{r}
# Create an empty dataFrame
aug_ind <- data.frame(missing_values = rep(NA, nrow(aug)))

# Count missing values
mss_list <- apply(aug[row.names(aug_ind), ], MARGIN = 1, FUN = function(x) {
  sum(x == "")
})
# Store missing values in the variables
aug_ind$missing_values <- mss_list

# Count inconsistencies
# No inconsistencies found

# Count number of outliers
# Cannot compute outliers for individuals, since there are categorical values in each row

# Check preliminary multivariant outliers for the dataset
mout.res <- Moutlier(aug[,c(3,13)], plot = FALSE) # The missing data are not imputed yet
llmout <- which(mout.res$md > mout.res$cutoff & mout.res$rd > mout.res$cutoff) # ids of mout
# Store the information about multivariant outliers
aug_ind$mvout <- 0
aug_ind$mvout[llmout] <- 1

# Summing up columns
aug_ind$sum_cols <- rowSums(aug_ind)

# Print the dataFrame
aug_ind
```

Now, we want to study which variables are correlated with this final column of summed up missing values, errors and outliers. Since it is a numerical variable, we can use condes() from FactoMineR library to do the profiling. The numerical variables associated with sum_cols are training_hours and city_development_index, but their correlation with the variable is pretty low. On the other hand, the most correlated categorical variables with the variable sum_cols are those with a p-value equals 0 and a high number of R2. Just for mentioning, the top three most correlated categorical variables are company_size (0.628), company_type (0.62) and education_level (0.33).

```{r}
# Add the column to the original dataset
aug$sum_cols <- aug_ind$sum_cols

# Profiling using condes
res.con <- condes(aug[2:ncol(aug)], 16) # We skip the first column of ids

# Correlated numerical variables
res.con$quanti

# Correlated categorical variables
res.con$quali

# The effect of the categories is not interested here, since there are still imputed data 
#res.con$category
```

In addition, we compute for every group of individuals the mean of missing values and then, rank the groups according to the computed mean.

```{r}
# Group means of f.city_development_index
sort(tapply(aug$sum_cols, aug$f.city_development_index, mean))
print("===========================================")
# Group means of gender
sort(tapply(aug$sum_cols, aug$gender, mean))
print("===========================================")
# Group means of relevent_experience
sort(tapply(aug$sum_cols, aug$relevent_experience, mean))
print("===========================================")
# Group means of enrolled_university
sort(tapply(aug$sum_cols, aug$enrolled_university, mean))
print("===========================================")
# Group means of education_level
sort(tapply(aug$sum_cols, aug$education_level, mean))
print("===========================================")
# Group means of major_discipline
sort(tapply(aug$sum_cols, aug$major_discipline, mean))
print("===========================================")
# Group means of f.training_hours
sort(tapply(aug$sum_cols, aug$f.training_hours, mean))
```

```{r}
# We remove this last column in the dataset, since we will not be working with it from now on
aug$sum_cols <- NULL
```


### 1.3. Imputation

### 1.3.1. Numeric Variables

Firstly, we need to assign NAs to the severe outliers found in numerical variables. The only numerical variables we have in our dataset are city_development_index and training_hours. Since we only have two variables and the former one does not have severe outliers, we will impute both severe and mild outliers using imputePCA().

Having a comparison of the quantiles of the imputed and unimputed numerical variables, we can state that the imputed values are correct, since the distributions of those two variables have not been greately altered.

```{r}
# Compute outliers
out_cdi <- calculate_outliers(aug$city_development_index, plot_yes = FALSE);
length(out_cdi$mild_indexes)
length(out_cdi$extreme_indexes)
out_th <- calculate_outliers(aug$training_hours, plot_yes = FALSE)
length(out_th$mild_indexes)
length(out_th$extreme_indexes)

# Assigning NAs
aug$city_development_index[out_cdi$mild_indexes] <- NA
aug$training_hours[out_th$mild_indexes] <- NA

# Imputation with imputePCA
res.pca <- imputePCA(aug[, c(3, 13)], ncp = 1)

# Validation
summary(res.pca$completeObs)
summary(aug[,c(3, 13)])

# Retain the imputed values
aug[,c(3, 13)] <- res.pca$completeObs
```

### 1.3.2. Factors

Firstly, we need to assign NAs to the empty strings "" in all categorical variables. Then, we will apply imputeMCA to impute categorical variables.

Having a comparison of the imputed and unimputed categorical variables, we can state that the imputed values are correct, since the frequencies of the categories in each of the variables do not seem to differ too much.

```{r}
# Assigning NAs
aug$gender[which(aug$gender == "")] <- NA
aug$enrolled_university[which(aug$enrolled_university == "")] <- NA
aug$education_level[which(aug$education_level == "")] <- NA
aug$major_discipline[which(aug$major_discipline == "")] <- NA
aug$experience[which(aug$experience == "")] <- NA
aug$company_size[which(aug$company_size == "")] <- NA
aug$company_type[which(aug$company_type == "")] <- NA
aug$last_new_job[which(aug$last_new_job == "")] <- NA

# Imputation with imputePCA (~ 25 seconds)
res.mca <- imputeMCA(aug[, c(2,4:12, 14)])

# Validation
summary(res.mca$completeObs)
print("====================================")
summary(aug[,c(2,4:12, 14)])

# Retain the imputed values
aug[,c(2,4:12, 14)] <- res.mca$completeObs
```

### 1.3.3. Removing Multivariate Outliers

Here, we use Moutlier() again to find multivariate outliers in the dataset and remove them from the dataset, since they can possibly decrease the performance of our final model.

```{r}
# Check multivariant outliers for the dataset
mout.res <- Moutlier(aug[,c(3,13)], plot = FALSE)
llmout <- which(mout.res$md > mout.res$cutoff & mout.res$rd > mout.res$cutoff) # ids of mout
length(llmout) # 484 multivariate outliers are found

# Comparing the values of the multivariate outliers
summary(aug[llmout,]) # Outliers
print("====================================")
summary(aug[-llmout,]) # No outliers

# Remove the multivariate outliers
aug <- aug[-llmout,]

# Reseting row names
row.names(aug) <- NULL
```

### 1.3.4. Final Data Cleaning Process

In this section we will prepare the data a little bit more in order to facilitate the later analysis. In the Univariate Descriptive Analysis section we found different problems of each variable, so here we are going to either apply different cleaning functions to the columns or just remove them.

```{r}
# We remove the column enrollee_id which do no provide any useful information
aug <- aug[,setdiff(names(aug),"enrollee_id")]

# Check remaining columns
names(aug)
```

Then, we should check if there are any duplicated observations in the dataset. Since the column enrollee_id is removed, now we can check that. We cannot do it previously because the column enrollee_id has the same number of unique values as the number of rows in the dataset. For that purpose, we can use the function duplicated() which returns all rows that appear at least twice in the dataset. It seems that there are 120 duplicated observations in our current dataset.

```{r}
# Find number of duplicated rows
sum(duplicated(aug)) # 120
# Find which observations are duplicated
duples <- which(duplicated(aug)) # One of the duplicated rows in the one in the position 247

# The function shows ids of rows that have already appeared before

# Show the duplicated observations of the observation 247
find_previous_duplicate <- function(df, row_id) {
  target <- df[row_id,]
  # Compare each row above row_id
  above_rows <- 1:(row_id - 1)
  for (i in above_rows) {
    row_i <- df[i,]
    # Compare element-wise, treating NA as equal
    equal <- mapply(function(a,b) a == b, row_i, target)
    if (all(equal)) return(i) # First match found
  }
  return(NA) # No previous match
}
previous_row_index <- find_previous_duplicate(aug, duples[1]) # Obs 247
aug[c(duples[1], previous_row_index),]

# Remove duplicated rows
aug <- aug[-duples,]

# Reseting row names
row.names(aug) <- NULL
```

The variables experience and company_size have too many levels, which will decrease the performance of the later modelling process. Therefore, we are going to group those levels into 4 groups at most.

```{r}
# Summary of experience
summary(aug$experience)

# Grouping
aug$f.experience[aug$experience %in% c("<1", "1", "2", "3", "4", "5")] <- "[<1,5]"
aug$f.experience[aug$experience %in% c("6", "7", "8", "9", "10")] <- "[6,10]"
aug$f.experience[aug$experience %in% c("11", "12", "13", "14", "15")] <- "[11,15]"
aug$f.experience[aug$experience %in% c("16", "17", "18", "19", "20", ">20")] <- "[16,>20]"

# Print
aug$f.experience
```

```{r}
# Summary of company_size
summary(aug$company_size)

# Small: <10, 10/49, 50-99
aug$f.company_size[aug$company_size %in% c("<10", "10/49", "50-99")] <- "Small"
# Medium: 100-500, 500-999
aug$f.company_size[aug$company_size %in% c("100-500", "500-999")] <- "Medium"
# Large: 1000-4999, 5000-9999
aug$f.company_size[aug$company_size %in% c("1000-4999", "5000-9999")] <- "Large"
# Very Large: 10000+
aug$f.company_size[aug$company_size == "10000+"] <- "Very-Large"

# Print
aug$f.company_size
```

### 1.3.5. Profiling on the Binary Target

Our target variable is (binary) categorical, so we will use catdes() from FactoMineR library to do the profiling. The most correlated categorical variables with the target variable target are city, f.city_development_index, company_size, experience, f.experience, enrolled_university, f.company_size and relevent_experience (p.value < 1e-50). On the other hand, the most correlated numerical variables (the only ones we have) are city_development_index (Eta2 = 0.11) and training_hours (Eta2 ~ 0).

According to the results in \$quanti, employees who have the intent of changing the job have the significant mean lower than the grand mean in the variables training_hours and city_development_index. Next, let's focus on the results given by \$category. The employees who have the intent of changing the job have low city_development_index and their current company's size is small. Their total experience in years is between less than a year and 5 years, so they do not have relevant experience, they were enrolled in a full time course at a university and their education level is a bachelor's. Based on this information, we can conclude that individuals seeking to change jobs are likely recent graduates with limited experience in the field.

```{r}
# Applying catdes on target
res.cat <- catdes(aug, num.var=13)

# Most significant categorical variables
res.cat$test.chi2

# Most significant numerical variables (the only two)
res.cat$quanti.var

# Description by the numerical variables
res.cat$quanti

# Description by the categorical variables
res.cat$category
```


















-> We still don't know if it is really necessary to create trian/test sets. Therefore, we are not going to apply the following cell yet.

Then, we split it into train and test sets. The test set will have 15% of the original dataset (2873), while the train set will have the remaining 85% (16285).

```{r}
# Split dataset
ts <- sample(1:nrow(aug), size = 0.15 * nrow(aug), replace = FALSE)
aug_test <- aug[ts,]
aug_train <- aug[-ts,]

# Get the dimension of the dataset
dim(aug) # 19158 obs + 14 variables
dim(aug_train) # 16285 obs + 14 variables
dim(aug_test) # 2873 obs + 14 variables
```